---
title: "6_Love_S9"
author: "Jeroen Gilis"
date: "07/12/2020"
output: html_document
---

This script (6_Love_S9.Rmd) is a seperate script for reproducing figure S9 from our publication. **In order to run this script (4_Love_DTU_subset.Rmd), the dataset Love_benchmark_datasets_scaledTPM.Rdata (or, alternatively, Love_benchmark_datasets_count.Rdata) is required.** This file can either be generated with the 1_Love_prepare.Rmd script or downloaded from Zenodo.

```{r setup, include=FALSE, echo=FALSE}
require("knitr")
opts_knit$set(root.dir = rprojroot::find_rstudio_root_file())
```

# Load libraries

```{r,message=FALSE}
library(ggplot2)
library(iCOBRA)
```

# Load DTU performance data

```{r}
# use scaledTPM results for the datasets based on Love et al. (and counts for all other datasets)
load(file="./Data/Love_DTU_results_scaledTPM.Rdata")

loveBenchmark <- c(
    loveDtuBenchmark_satuRn,
    loveDtuBenchmark_DoubleExpSeq,
    loveDtuBenchmark_limmaDiffsplice,
    loveDtuBenchmark_edgeRdiffsplice,
    loveDtuBenchmark_DEXSeq,
    loveDtuBenchmark_DRIMSeq,
    loveDtuBenchmark_NBSplice
)

### Remove empty enteries (due to not tested - aka to many samples for reasonable runtime)
loveBenchmark  <- loveBenchmark[which(
    sapply(loveBenchmark, function(x) ! is.null(x$dtuAnalysis))
)]

rm(loveDtuBenchmark_satuRn,
    loveDtuBenchmark_DoubleExpSeq,
    loveDtuBenchmark_limmaDiffsplice,
    loveDtuBenchmark_edgeRdiffsplice,
    loveDtuBenchmark_DEXSeq,
    loveDtuBenchmark_DRIMSeq,
    loveDtuBenchmark_NBSplice)
invisible(gc())
```

# Load and wrangle truth file

```{r}
# define the truth (copy from swimming downstream paper by Love et al.,F1000Research 2018, 7:952)
load("./Data/Love_metadata.rda")
txdf <- txdf[match(rownames(tpms), txdf$TXNAME),]
txdf$dtu.genes <- iso.dtu | iso.dte & !iso.dte.only
full.dtu.genes <- unique(txdf$GENEID[txdf$dtu.genes])

txp.exprs <- rowSums(tpms) > 0
dtu.dte.genes <- unique(txdf$GENEID[iso.dte & !iso.dte.only])
txdf$full.dtu <- iso.dtu | (txdf$GENEID %in% dtu.dte.genes & txp.exprs)
dtu.txps <- txdf$TXNAME[txdf$full.dtu]
```

# Figure S9A

```{r}
tval <- loveBenchmark[["satuRn_samples_used_6_rep_2_filterLenient"]]$dtuAnalysis$tvalue
pval <- loveBenchmark[["satuRn_samples_used_6_rep_2_filterLenient"]]$dtuAnalysis$p_value_raw
  
zval <- qnorm(pval/2) * sign(tval)
zval_mid <- zval[abs(zval) < 10]
zval_mid <- zval_mid[!is.na(zval_mid)]

png("./Results/Love_benchmark/figS9A.png",
    width     = 3.5,
    height    = 2.5,
    units     = "in",
    res       = 300,
    pointsize = 6) # start export
plot <- locfdr::locfdr(zval_mid, main = "Love dataset - 6v6 - edgeR filter - repeat 2")
dev.off()
```

# Figure S9B

```{r}
Love_select <- loveBenchmark[["satuRn_samples_used_6_rep_2_filterLenient"]]$dtuAnalysis
    
pvalues <- as.data.frame(cbind(Love_select$p_value_raw,Love_select$p_value))
colnames(pvalues) <- c("theoretical_null", "empirical_null")
rownames(pvalues) <- rownames(Love_select)
    
# generate truth file
truth_full_new <- as.data.frame(cbind(rownames(pvalues),rep(0,nrow(pvalues))))
colnames(truth_full_new) <- c("TXNAME","truth")
rownames(truth_full_new) <- truth_full_new$TXNAME
truth_full_new$truth <- as.numeric(as.character(truth_full_new$truth))
truth_full_new[which(sub("\\_.*", "", truth_full_new$TXNAME) %in% dtu.txps),"truth"] <- 1

cobra <- COBRAData(pval = pvalues, truth = truth_full_new)
cobra <- calculate_adjp(cobra)
cobra1perf <- calculate_performance(cobra, binary_truth = "truth", cont_truth = "none", splv = "none", aspects = c("fdrtpr", "fdrtprcurve", "overlap"))

cobraplot <- prepare_data_for_plot(cobra1perf, colorscheme = "Dark2", facetted = TRUE)

new_col <- c(rep(c("black", "grey64", "#999999"),4),rep("white",6))
names(new_col) <- names(cobraplot@plotcolors)
cobraplot@plotcolors <- new_col
    
plot <- plot_fdrtprcurve(cobraplot, xaxisrange = c(0, 0.4), yaxisrange = c(0,1))

plot <- plot + 
    ggtitle(label="Love dataset - 6v6 - edgeR filter - repeat 2") +
    xlab("FDP") +
    theme(strip.background = element_blank(),
    strip.text.x = element_blank(),
    plot.title = element_text(size=10),
    axis.text.x = element_text(size = 10),
    axis.text.y = element_text(size = 10),
    axis.title.x = element_text(size = 10),
    axis.title.y = element_text(size = 10))

png("./Results/Love_benchmark/figS9B.png",
    width = 5.5,
    height = 4.5,
    units = "in",
    res = 200,
    pointsize = 4) # start export
print(plot)
dev.off()
```

