---
title: "satuRn: Effect of the magnitude of the DTU signal on performance"
author: "Jeroen Gilis"
date: "09/07/2022"
output:
  html_document:
    code_download: yes
    theme: cosmo
    toc: yes
    toc_float: yes
    highlight: tango
    number_sections: yes
  pdf_document:
    toc: yes
---

In our real bulk and single-cell benchmark studies, DTU between groups of 
samples or cells is introduced by starting with a mock comparison (no DTU 
signal) and subsequently swapping 10% of the  transcript counts 
(within the same gene) within 1 group of samples or cells. As such, the swapping
strategy will induce DTU signal of different magnitudes. We followed the 
reviewer’s suggestion of assessing the performance of the different methods as a
function of the strength of the induced signal.

Specifically, we have included four additional figures in the Extended data 
(**Figures S16-S19**), corresponding to the four benchmarks that used swapping, 
where we stratify FDR-TPR curves based on the strength of the induced DTU 
signal. As an estimate for the strength of the signal, we take the difference in
the empirical average transcript usage between the two groups of cells/samples. 
The empirical usage was computed by dividing the transcript-level count with the
corresponding gene-level count in each cell/sample. Based on this metric, the 
FDR-TPR curves are stratified in 6 categories: a difference in transcript usage 
between groups of 0-10%, 10-20%, 20-30%, 30-40%, 40-50% and 50-100%. A brief 
discussion of the results is included in the "Performance on real single-cell 
data paragraph" of the Results section of the manuscript.

*First*, as expected, the ability of all methods to detect DTU decreases when 
the strength of the DTU signal decreases. *Second*, the difference of transcript 
usage that can be correctly detected is dependent on the data: small differences 
in transcript usage can be more easily detected in bulk data as compared to 
single-cell data, likely owing to the larger sequencing depth and lower 
variability of bulk datasets. *Third*, inference improves for all strata when 
the number of cells or samples increases. *Fourth*, we observed in all benchmark
datasets that satuRn and DoubleExpSeq are more successful in detecting small 
differences as compared to the other methods. As such, the performance gain of 
satuRn and DoubleExpSeq in the overall results (without stratification) can be
attributed to picking up DTU with smaller effect sizes. As a final note, having
different transcript-level filtering procedures (edgeR filter versus DRIMSeq 
filter) did not affect these findings. Adopting scaledTPM abundances instead of 
raw counts did not affect the results, either (not shown, but can be obtained
by running this script and changing input data from counts to scaledTPM).


```{r setup, include=FALSE, echo=FALSE}
require("knitr")
opts_knit$set(root.dir = rprojroot::find_rstudio_root_file())
```

# Load libraries

```{r,message=FALSE,warning=FALSE}
library(ggplot2)
library(dplyr)
library(iCOBRA)
```

# Helper functions

## get_pvalues

To get pvalues for the different methods on the different datasets

```{r}
get_pvalues <- function(dataset,nrep,nmethods) {
  
  get_all_txs <- function(dataset,nrep){
  
    all_txs <- c()
    for (i in 1:nrep) {
      current_txs <- as.character(dataset[[i]]$dtuAnalysis$TXNAME)
      all_txs <- c(all_txs, current_txs)
      all_txs <- unique(all_txs)
    }
    return(all_txs)
  }

  all_txs <- get_all_txs(dataset,nrep)
  pvalues <- matrix(data=NA, nrow=length(all_txs), ncol=nmethods*nrep)
  rownames(pvalues) <- all_txs
  
  for (i in 1:(nmethods*nrep)) {
    pvalues_new <- data.frame(dataset[[i]]$dtuAnalysis$p_value)
    rownames(pvalues_new) <- as.character(dataset[[i]]$dtuAnalysis$TXNAME)
    colnames(pvalues_new) <- names(dataset)[i]

    pvalues[,i] <- pvalues_new[match(rownames(pvalues),rownames(pvalues_new)),1]
  }
  
  pvalues <- apply(pvalues, 2, as.numeric)
  rownames(pvalues) <- all_txs
  pvalues <- as.data.frame(pvalues)
  pvalues <- pvalues[order(rownames(pvalues)),]
  
  if(nmethods == 6){
    colnames(pvalues) <- c("satuRn", "limma_diffsplice", "DEXSeq", "DRIMSeq","edgeR_diffsplice", "DoubleExpSeq")
  } else if(nmethods == 4) {
    colnames(pvalues) <- c("satuRn", "limma_diffsplice","edgeR_diffsplice", "DoubleExpSeq")
  } else if(nmethods == 7){
    colnames(pvalues) <- c("satuRn", "limma_diffsplice", "DEXSeq", "DRIMSeq","edgeR_diffsplice", "DoubleExpSeq", "NBSplice")
  }
  return(pvalues)
}
```

## .get_MeanDiff

```{r}
.get_MeanDiff <- function(counts, tx2gene, txSwapped, quantiles, diagplot = TRUE){
  
    otherCounts <- satuRn:::.getOtherCount(counts, tx2gene)
    totalCounts <- counts + otherCounts
    usage <- counts/totalCounts
    
    group_a <- txSwapped[[1]]$design[txSwapped[[1]]$design$condition == "a", "sample_id"]
    group_b <- txSwapped[[1]]$design[txSwapped[[1]]$design$condition == "b", "sample_id"]
    
    mean_usage_a <- rowMeans(usage[,group_a], na.rm=TRUE)
    mean_usage_b <- rowMeans(usage[,group_b], na.rm=TRUE)
    
    diff_usage <- abs(mean_usage_a - mean_usage_b)
    names(diff_usage) <- rownames(usage)
    
    diff_usage <- diff_usage[match(txSwapped[[1]]$metaInfo$TXNAME,
                                   names(diff_usage))]
    
    diff_shift <- diff_usage[txSwapped[[1]]$metaInfo$txSwapped == TRUE]
    diff_null <- diff_usage[txSwapped[[1]]$metaInfo$txSwapped == FALSE]
    cutoffs <- quantile(diff_null, quantiles)
    
    if(diagplot == TRUE){
        hist(diff_null, breaks=40, col=rgb(0,1,0,0), xlim=c(0,1), main = "Difference in mean usage",freq = F)
        hist(diff_shift, breaks=40, add=TRUE, col=rgb(0,0,1,0.7),freq = F)
        hist(diff_null, breaks=40, col=rgb(0,1,0,0.5), add=TRUE, freq=F)
        abline(v=cutoffs, col="grey50", lty=2)      
    }
    
    return(diff_usage)
}
```

## .fdrtpr_helper

```{r}
.fdrtpr_helper <- function(pvalues, truth, splv_col, cutoffs, j, diagplot, sampleSize){
    
  truth <- truth %>% 
      mutate(category = cut(truth[,splv_col],
                            breaks = cutoffs))
  
  if(diagplot == TRUE){
    hist(truth$MeanDiffs[truth$gene_modified == 0], 
         xlim=c(0,1), breaks=40, freq=FALSE, col=rgb(0,1,0,0.5))
    hist(truth$MeanDiffs[truth$gene_modified == 1], 
         breaks=40, freq=FALSE, col=rgb(0,0,1,0.5), add=TRUE)
    abline(v=cutoffs, lty=2, col="grey50")
    
    hist(truth$MeanDiffs[truth$gene_modified == 0], 
         xlim=c(0,1), breaks=40, col=rgb(0,1,0,0.5))
    hist(truth$MeanDiffs[truth$gene_modified == 1], 
         breaks=40, col=rgb(0,0,1,0.5), add=TRUE)
    abline(v=cutoffs, lty=2, col="grey50")
  }
  
  cobra <- COBRAData(pval = pvalues, truth = truth)
  cobra <- calculate_adjp(cobra)
  
  cobra1perf <- calculate_performance(cobra, 
                                      binary_truth = "gene_modified", 
                                      cont_truth = "none", 
                                      splv = "category",
                                      maxsplit = Inf,
                                      aspects = c("fdrtpr", "fdrtprcurve", "overlap"))
  
  cobraplot <- prepare_data_for_plot(cobra1perf, 
                                     colorscheme = "Dark2", 
                                     facetted = TRUE)
  
  if(ncol(pvalues) == 6){
    
    colors <- c("#E69F00", "#56B4E9", "#009E73", "#F0E442", "#0072B2", "black", "#999999")
    whites <- length(which(cobraplot@plotcolors == "white"))
    notwhites <- length(which(cobraplot@plotcolors != "white"))/length(colors)
    
    new_col <- c(rep(colors,notwhites),
                rep("white",whites))
    names(new_col) <- names(cobraplot@plotcolors)
    
    current_title <- paste0(sampleSize, " versus ", sampleSize, " - edgeR filter - count")
  } else if(ncol(pvalues) == 4){
    
    colors <- c("#56B4E9", "#F0E442", "#0072B2", "black", "#999999")
    whites <- length(which(cobraplot@plotcolors == "white"))
    notwhites <- length(which(cobraplot@plotcolors != "white"))/length(colors)
    
    new_col <-c(rep(colors,notwhites),
                rep("white", whites))
    names(new_col) <- names(cobraplot@plotcolors)

    current_title <- paste0(sampleSize, " versus ", sampleSize, " - edgeR filter - count")
  } else if(ncol(pvalues) == 7){
    colors <- c("#E69F00", "#56B4E9", "#009E73", "#F0E442", "#0072B2", "#D55E00", "black", "#999999")
    whites <- length(which(cobraplot@plotcolors == "white"))
    notwhites <- length(which(cobraplot@plotcolors != "white"))/length(colors)
    
    new_col <-c(rep(colors,notwhites),
                rep("white", whites))
    names(new_col) <- names(cobraplot@plotcolors)

    current_title <- paste0(sampleSize, " versus ", sampleSize, " - edgeR filter - count")
  }
  
  cobraplot@plotcolors <- new_col
  plot <- plot_fdrtprcurve(cobraplot, 
                           xaxisrange = c(0, 0.4), 
                           yaxisrange = c(0,1),
                           pointsize = 2,
                           linewidth = 0.75)
  plot <- plot + scale_color_manual(values = new_col, limits=force)
  
  plot <- plot + 
      ggtitle(current_title) +
      theme(strip.text.x = element_text(size = 10),
            plot.title = element_text(size=12,face="bold"),
            axis.text.x = element_text(size = 10),
            axis.text.y = element_text(size = 10),
            axis.title.x = element_text(size = 10),
            axis.title.y = element_text(size = 10)) +
     facet_wrap(~splitval, ncol = 3)
  
  return(plot)
}
```

## visualize_datasets

```{r}
visualize_datasets <- function(dataset,
                               saveName,
                               metaInfo,
                               nFilters,
                               nSampleSizes,
                               nRep,
                               nmethods,
                               selection,
                               selection2) {

    count <- 0
    nPlots <- nFilters * nSampleSizes
    print(paste("This function will generate",  nPlots, "plots"))
  
    for (j in 1:nPlots) {
        print(j)
        sampleSize <- gsub("([0-9]+).*$", "\\1", selection2[j])
        Chen_current_outer <- dataset[grepl(selection[j], names(dataset))]
        metaInfo_current_outer <- metaInfo[grepl(selection2[j], names(metaInfo))]

        resList <- c()
        count <- count+1

        for (k in 1:length(nRep)) {
            Chen_current_inner <- Chen_current_outer[grepl(nRep[k], names(Chen_current_outer))]

            metaInfo_current_inner <- metaInfo_current_outer[grepl(nRep[k], names(metaInfo_current_outer))]
    
            pvalues <- get_pvalues(Chen_current_inner, nrep=1, nmethods)

            truth_file <- Chen_current_inner[[1]]$dtuAnalysis
    
            if(count <= nSampleSizes){
                txSwapped <- metaInfo_current_inner[1]          
                } else {
                txSwapped <- metaInfo_current_inner[2]
            }
    
            truth_file <- truth_file[,c("TXNAME","p_value")]
            
            ################### extra part to compute splv's ###################
            counts <- txSwapped[[1]]$data
            tx2gene <- txSwapped[[1]]$metaInfo[,c(1,2)]
            colnames(tx2gene) <- c("isoform_id", "gene_id")
            
            #### Means
            quantiles <- c(seq(0.5,0.9,0.1),0.95) # quantiles on diagplot
            MeanDiffs <- .get_MeanDiff(counts, tx2gene, txSwapped, quantiles, diagplot=TRUE)
            
            ####################################################################
    
            txSwapped[[1]]$metaInfo <- txSwapped[[1]]$metaInfo[match(truth_file$TXNAME,txSwapped[[1]]$metaInfo$TXNAME),]
    
            truth_file$gene_modified <- as.numeric(txSwapped[[1]]$metaInfo$txSwapped)
            truth_file <- truth_file[order(truth_file$TXNAME),]
            truth_file$MeanDiffs <- MeanDiffs[match(truth_file$TXNAME, names(MeanDiffs))]
            rownames(truth_file) <- truth_file$TXNAME

            resList[[k]] <- list(pvalues,truth_file)
        }
        
        rownames(resList[[1]][[1]]) <- paste0(rownames(resList[[1]][[1]]), "_1")
        rownames(resList[[2]][[1]]) <- paste0(rownames(resList[[2]][[1]]), "_2")
        rownames(resList[[3]][[1]]) <- paste0(rownames(resList[[3]][[1]]), "_3")
        
        pvalues_full <- rbind(resList[[1]][[1]],resList[[2]][[1]],resList[[3]][[1]])
        
        rownames(resList[[1]][[2]]) <- paste0(rownames(resList[[1]][[2]]), "_1")
        rownames(resList[[2]][[2]]) <- paste0(rownames(resList[[2]][[2]]), "_2")
        rownames(resList[[3]][[2]]) <- paste0(rownames(resList[[3]][[2]]), "_3")
        
        truth_file_full <- rbind(resList[[1]][[2]],resList[[2]][[2]],resList[[3]][[2]])
        
        plot <- .fdrtpr_helper(pvalues = pvalues_full, 
                               truth = truth_file_full, 
                               splv_col = 4, # the column on which to stratify
                               cutoffs = c(seq(0,0.4,0.1),1),
                               j = j,
                               sampleSize = sampleSize,
                               diagplot = TRUE)

        png(paste("./Results/Benchmarks_additional/DTU_strength/", saveName,
                  "_", selection2[j], "_", sub("^[^f]*", "", selection[j]),
                  "_counts.png", sep = ""),
                    width     = 8,
                    height    = 6,
                    units     = "in",
                    res       = 200,
                    pointsize = 4) # start export
        print(plot)
        dev.off()
       
  }
}
```

# Tasic dataset

# Load truth file

```{r}
load(file="./Data/Tasic_benchmark_datasets_count.Rdata")
#load(file="./Data/Tasic_benchmark_datasets_scaledTPM.Rdata")
metaInfo <- c(TasicBenchmarkLenient,TasicBenchmarkStringent)
rm(TasicBenchmarkLenient,TasicBenchmarkStringent)
```

# Load DTU performance data

```{r}
load(file="./Data/Tasic_DTU_results_count.Rdata")
#load(file="./Data/Tasic_DTU_results_scaledTPM.Rdata")
TasicBenchmark <- c(
    TasicDtuBenchmark_satuRn,
    TasicDtuBenchmark_limmaDiffsplice,
    TasicDtuBenchmark_DEXSeq,
    TasicDtuBenchmark_DRIMSeq,
    TasicDtuBenchmark_edgeRdiffsplice,
    TasicDtuBenchmark_DoubleExpSeq)

### Remove empty entries (due to not tested - aka to many samples for reasonable runtime)
TasicBenchmark  <- TasicBenchmark[which(
    sapply(TasicBenchmark, function(x) ! is.null(x$dtuAnalysis)))]

rm(TasicDtuBenchmark_satuRn,
    TasicDtuBenchmark_limmaDiffsplice,
    TasicDtuBenchmark_DEXSeq,
    TasicDtuBenchmark_DRIMSeq,
    TasicDtuBenchmark_edgeRdiffsplice,
    TasicDtuBenchmark_DoubleExpSeq)
invisible(gc())
```

## Generate FDR-TPR curves

```{r}
selection <- c("20_rep_._filterLenient")
selection2 <- c("20_rep")

visualize_datasets(dataset=TasicBenchmark,
                   saveName = "Tasic",
                   metaInfo=metaInfo,
                   nFilters=1,
                   nSampleSizes=1,
                   nRep=c("rep_1","rep_2","rep_3"),
                   nmethods=6,
                   selection = selection,
                   selection2=selection2)

selection <- c("_75_rep_._filterLenient","200_rep_._filterLenient")
selection2 <- c("75_rep","200_rep")

visualize_datasets(dataset=TasicBenchmark,
                   saveName = "Tasic",
                   metaInfo=metaInfo,
                   nFilters=1,
                   nSampleSizes=2,
                   nRep=c("rep_1","rep_2","rep_3"),
                   nmethods=4,
                   selection = selection,
                   selection2=selection2)
```

# Chen dataset

## Load truth file

```{r}
load(file="./Data/Chen_benchmark_datasets_count.Rdata")
metaInfo <- c(ChenBenchmarkLenient,ChenBenchmarkStringent)
rm(ChenBenchmarkLenient,ChenBenchmarkStringent)
```

## Load DTU performance data

```{r}
load(file="./Data/Chen_DTU_results_count.Rdata")

ChenBenchmark <- c(
    ChenDtuBenchmark_satuRn,
    ChenDtuBenchmark_limmaDiffsplice,
    ChenDtuBenchmark_DEXSeq,
    ChenDtuBenchmark_DRIMSeq,
    ChenDtuBenchmark_edgeRdiffsplice,
    ChenDtuBenchmark_DoubleExpSeq)

### Remove empty entries (due to not tested - aka to many samples for reasonable runtime)
ChenBenchmark  <- ChenBenchmark[which(
    sapply(ChenBenchmark, function(x) ! is.null(x$dtuAnalysis)))]

rm(ChenDtuBenchmark_satuRn,
    ChenDtuBenchmark_limmaDiffsplice,
    ChenDtuBenchmark_DEXSeq,
    ChenDtuBenchmark_DRIMSeq,
    ChenDtuBenchmark_edgeRdiffsplice,
    ChenDtuBenchmark_DoubleExpSeq)
invisible(gc())
```

## Generate FDR-TPR curves

```{r}
selection <- c("20_rep_._filterLenient")
selection2 <- c("20_rep")
visualize_datasets(dataset=ChenBenchmark,
                   saveName = "Chen",
                   metaInfo=metaInfo,
                   nFilters=1,
                   nSampleSizes= 1,
                   nRep=c("rep_1","rep_2","rep_3"),
                   nmethods=6,
                   selection = selection,
                   selection2=selection2)

selection <- c("_50_rep_._filterLenient")
selection2 <- c("50_rep")
visualize_datasets(dataset=ChenBenchmark,
                   saveName = "Chen",
                   metaInfo=metaInfo,
                   nFilters=1,
                   nSampleSizes=1,
                   nRep=c("rep_1","rep_2","rep_3"),
                   nmethods=4,
                   selection = selection,
                   selection2=selection2)
```

# Darmanis dataset

# Load truth file

```{r}
load(file="./Data/Darmanis_benchmark_datasets_count.Rdata")
metaInfo <- c(DarmanisBenchmarkLenient,DarmanisBenchmarkStringent)
rm(DarmanisBenchmarkLenient,DarmanisBenchmarkStringent)
```

# Load DTU performance data

```{r}
load(file="./Data/Darmanis_DTU_results_count.Rdata")

DarmanisBenchmark <- c(
    DarmanisDtuBenchmark_satuRn,
    DarmanisDtuBenchmark_limmaDiffsplice,
    DarmanisDtuBenchmark_DEXSeq,
    DarmanisDtuBenchmark_DRIMSeq,
    DarmanisDtuBenchmark_edgeRdiffsplice,
    DarmanisDtuBenchmark_DoubleExpSeq)

### Remove empty entries (due to not tested - aka to many samples for reasonable runtime)
DarmanisBenchmark  <- DarmanisBenchmark[which(
    sapply(DarmanisBenchmark, function(x) ! is.null(x$dtuAnalysis)))]

rm(DarmanisDtuBenchmark_satuRn,
    DarmanisDtuBenchmark_limmaDiffsplice,
    DarmanisDtuBenchmark_DEXSeq,
    DarmanisDtuBenchmark_DRIMSeq,
    DarmanisDtuBenchmark_edgeRdiffsplice,
    DarmanisDtuBenchmark_DoubleExpSeq)
invisible(gc())
```

# Generate FDR-TPR curves

```{r}
selection <- c("20_rep_._filterLenient")
selection2 <- c("20_rep")

visualize_datasets(dataset=DarmanisBenchmark,
                   saveName = "Darmanis",
                   metaInfo=metaInfo,
                   nFilters=1,
                   nSampleSizes=1,
                   nRep=c("rep_1","rep_2","rep_3"),
                   nmethods=6,
                   selection = selection,
                   selection2=selection2)

selection <- c("_50_rep_._filterLenient","100_rep_._filterLenient")
selection2 <- c("50_rep","100_rep")

visualize_datasets(dataset=DarmanisBenchmark,
                   saveName = "Darmanis",
                   metaInfo=metaInfo,
                   nFilters=1,
                   nSampleSizes=2,
                   nRep=c("rep_1","rep_2","rep_3"),
                   nmethods=4,
                   selection = selection,
                   selection2=selection2)
```

# Load truth file

```{r}
load(file="./Data/GTEx_benchmark_datasets_count.Rdata")
metaInfo <- c(gtexBenchmarkDataLenient,gtexBenchmarkDataStringent)
rm(gtexBenchmarkDataLenient,gtexBenchmarkDataStringent)
```

# Load DTU performance data

```{r}
load(file="./Data/GTEx_DTU_results_count.Rdata")

gtexBenchmark <- c(
    gtexDtuBenchmark_satuRn,
    gtexDtuBenchmark_limmaDiffsplice,
    gtexDtuBenchmark_DEXSeq,
    gtexDtuBenchmark_DRIMSeq,
    gtexDtuBenchmark_edgeRdiffsplice,
    gtexDtuBenchmark_DoubleExpSeq,
    gtexDtuBenchmark_NBSplice
)

### Remove empty entries (due to not tested due to too many samples for reasonable runtime)
gtexBenchmark  <- gtexBenchmark[which(
    sapply(gtexBenchmark, function(x) ! is.null(x$dtuAnalysis))
)]

rm(gtexDtuBenchmark_satuRn,
    gtexDtuBenchmark_limmaDiffsplice,
    gtexDtuBenchmark_DEXSeq,
    gtexDtuBenchmark_DRIMSeq,
    gtexDtuBenchmark_edgeRdiffsplice,
    gtexDtuBenchmark_DoubleExpSeq,
    gtexDtuBenchmark_NBSplice)
invisible(gc())
```

# Generate FDR-TPR curves

```{r}
selection <- c("5_rep_.(_|0_)filterLenient","20_rep_.(_|0_)filterLenient")
selection2 <- c("5_rep", "20_rep")

visualize_datasets(dataset=gtexBenchmark,
                   saveName="GTEx",
                   metaInfo=metaInfo,
                   nFilters=1,
                   nSampleSizes=2,
                   nRep=c("rep_1","rep_2","rep_3"),
                   nmethods=7,
                   selection = selection,
                   selection2 = selection2)

selection <- c("50_rep_.(_|0_)filterLenient")
selection2 <- c("50_rep")

visualize_datasets(dataset=gtexBenchmark,
                   saveName="GTEx",
                   metaInfo=metaInfo,
                   nFilters=1,
                   nSampleSizes=1,
                   nRep=c("rep_1","rep_2","rep_3"),
                   nmethods=4,
                   selection = selection,
                   selection2 = selection2)
```


