---
title: "satuRn: Effect of filtering and abundance measure on performance"
author: "Jeroen Gilis"
date: "08/07/2022"
output:
  html_document:
    code_download: yes
    theme: cosmo
    toc: yes
    toc_float: yes
    highlight: tango
    number_sections: yes
  pdf_document:
    toc: yes
---

We have followed the reviewer’s suggestion of including simplified figures 
that compare the different filtering and counting approaches on a single plot 
per tool. These figures have been included as `Extended data Figures S10-S13`
and are discussed in the “Performance on simulated bulk RNA-seq datasets” 
paragraph in the Results section of the manuscript.

These figures show more clearly that neither sample size nor the filtering 
strategy had a profound impact on the ranking of the performances of the 
different DTU methods; satuRn, DEXSeq and DoubleExpSeq remain the best 
performing methods overall. However, we did find that the effects of filtering 
and counting differ between the different datasets.

For the Love et al. simulated bulk dataset, filtering more stringently strongly
improves performance. In addition, both performance and FDR control are much 
better when using scaledTPM abundances, as compared to using counts. The strong 
positive effect of using scaledTPM abundances are only observed for this 
dataset, which we further discuss in our manuscript. 

For the simulated bulk datasets by Van den Berge et al., we also observe a 
positive effect of stringent filtering. The difference between scaledTPM and 
raw count abundances is negligible.

For the real datasets, a quite different pattern is observed. For the GTEx 
bulk dataset and the single-cell datasets by Tasic et al. and Chen et al., 
the effects of filtering are limited and using counts performs slightly better 
than using scaledTPM abundances.

Finally, stringent filtering had a strong positive impact on performance on the
Darmanis et al. dataset. Note that this dataset is by far the sparsest dataset 
in our benchmarks (Extended data Figure S32 and Table S1). The differences in 
performance between scaledTPM and raw count abundances are small.

Taken together, using scaledTPM abundances is only beneficial in the simulations
by Love et al., for reasons we explain in below. Therefore, we recommend using 
raw counts abundances. With respect to filtering, the more stringent DRIMSeq 
filtering strategy only seems beneficial in simulated data and in very sparse 
single-cell data. Thus, for bulk data and regular single-cell data the lenient 
edgeR filtering strategy seems sufficient to increase the signal-to-noise ratio.

```{r setup, include=FALSE, echo=FALSE}
require("knitr")
opts_knit$set(root.dir = rprojroot::find_rstudio_root_file())
```

# Load libraries

```{r,message=FALSE,warning=FALSE}
library(ggplot2)
library(iCOBRA)
library(stringr)
library(purrr)
library(gridExtra)
```

# Helper functions

## computePlot

```{r}
computePlot <- function(result, metadata, print = FALSE){
  
  # loop over independent datasets (repeats must be joined)
  names_norep <- str_replace(names(result), 
                             pattern = "_rep_\\d_", 
                             replacement = "_")
  names_norep <- levels(as.factor(names_norep))
  
  # store resulting plots
  plotlist <- list()
  
  for (analysis in seq_along(names_norep)) {
    
    # select the current datasets
    current_name <- names_norep[analysis]
    matches <- which(str_replace(names(result), 
                                 pattern = "_rep_\\d_", 
                                 replacement = "_") 
                     == current_name)
    current_result <- result[matches]
    matches <- which(str_replace(names(metadata), 
                                 pattern = "_rep_\\d_", 
                                 replacement = "_") 
                     == gsub(".*samples", "samples", current_name))
    current_metadata <- metadata[matches]
    
    # Get relevant results (TXNAME and p_value) from nested list
    # Append results from the repeats for correct FDR-TPR calculations
    # TODO: how to do this with purrr::map_depth()
    pvalues <- c()
    TXNAMES_res <- c()
    TXNAMES_meta <- c()
    txSwapped <- c()
    for (i in seq_along(current_result)) {
      pvalues <- c(pvalues, current_result[[i]]$dtuAnalysis$p_value)
      TXNAMES_res <- c(TXNAMES_res, paste0(current_result[[i]]$dtuAnalysis$TXNAME, "_", i))
      TXNAMES_meta <- c(TXNAMES_meta, paste0(current_metadata[[i]]$metaInfo$TXNAME, "_", i))
      txSwapped <- c(txSwapped, current_metadata[[i]]$metaInfo$txSwapped)
    }
    
    # match results with ground truth
    txSwapped <- txSwapped[match(TXNAMES_res,TXNAMES_meta)]
    
    # wrangle for iCOBRA input
    pval_cobra <- as.data.frame(cbind(pvalues))
    rownames(pval_cobra) <- TXNAMES_res
    truth_cobra <- as.data.frame(cbind(txSwapped))
    rownames(truth_cobra) <- TXNAMES_res
    if(max(truth_cobra$txSwapped) == 2){
      truth_cobra$txSwapped <- ifelse(truth_cobra$txSwapped == 1, 0, 1)
    }
    if(is.character(truth_cobra$txSwapped)){
      truth_cobra$txSwapped <- as.factor(as.numeric(truth_cobra$txSwapped))
    }
    
    # create sensible "method" name
    quant <- tail(str_split(current_name, "_")[[1]], n=1)
    filter <- tail(str_split(current_name, "_")[[1]], n=2)[1]
    filter <- tolower(gsub("filter", "", filter))
    colnames(pval_cobra) <- paste0(quant, "_", filter)
    
    # use iCOBRA to create plot
    cobra <- COBRAData(pval = pval_cobra, truth = truth_cobra)
    cobra <- calculate_adjp(cobra)
    cobra1perf <- suppressMessages(
      calculate_performance(cobra, 
                            binary_truth = "txSwapped", 
                            cont_truth = "none", 
                            splv = "none", 
                            aspects = c("fdrtpr", "fdrtprcurve", "overlap"))
    )
    cobraplot <- prepare_data_for_plot(cobra1perf, 
                                       colorscheme = "Dark2", 
                                       facetted = TRUE)
    
    if(print == TRUE){
      pp <- plot_fdrtprcurve(cobraplot)
      print(pp)
    }
    
    # store plot
    plotlist[[analysis]] <- cobraplot
  }
  
  names(plotlist) <- names_norep
  return(plotlist)
}
```

## makePlot

```{r}
makePlot <- function(plotlist, 
                     colors,
                     dataset,
                     type = "cells",
                     xaxisrange = c(0, 0.4),
                     yaxisrange = c(0, 1),
                     linewidth = 1,
                     pointsize = 3,
                     stripsize = 15,
                     titlecol = "black"){
  
  plot_data_lines <- list()
  plot_data_points <- list()
  
  for (i in seq_along(plotlist)) {
    plot_data_lines[[i]] <- fdrtprcurve(plotlist[[i]])
    plot_data_points[[i]] <- fdrtpr(plotlist[[i]])
  }
  
  plot_data_lines <- do.call(rbind, plot_data_lines)
  plot_data_points <- do.call(rbind, plot_data_points)
  
  xasp <- "FDR"
  yasp <- "TPR"
  
  thresholds <- sort(unique(as.numeric(gsub("thr", "", plot_data_points$thr))))
  plot_data_points$method2.satis <- paste0(plot_data_points$method,
                                           plot_data_points$satis)
  
  if (!(isTRUE(facetted(plotlist[[1]])))) {
    plot_data_points$method <- plot_data_points$fullmethod
    plot_data_lines$method <- plot_data_lines$fullmethod
    plot_data_points$method2.satis <- paste0(plot_data_points$fullmethod,
                                             plot_data_points$satis)
  }
  
  ## Number of colors/linetypes
  nlevs <- length(unique(plot_data_lines$method))
  ## Number of thresholds
  nthr <- length(unique(plot_data_points$thr))
  
  # Get desired colors
  newcol <- c(plotcolors(plotlist[[1]]),
              plotcolors(plotlist[[2]]),
              plotcolors(plotlist[[3]]),
              plotcolors(plotlist[[4]]))
  newcol[c(1,5,13,17,25,29,37,41)]<- rep(colors,  each = 2)
  
  # construct title
  method <- gsub("_samples.*", "", names(plotlist)[1])
  nCells <- str_extract(names(plotlist)[1], "[[:digit:]]+")
  title <- paste0(dataset, ": ", nCells, " ", type, " (", method, ")")
  
  pp <- ggplot(plot_data_lines, 
               aes_string(x = xasp, y = yasp, group = "method", colour = "method")) + 
    geom_vline(xintercept = seq(0, xaxisrange[2], 0.1), 
               colour = "lightgrey", linetype = "dashed") + 
    geom_vline(xintercept = thresholds, linetype = "dashed") + 
    geom_path(size = linewidth, aes_string(linetype = "method")) + 
    scale_linetype_manual(values = rep("solid", nlevs), guide = "none") + 
    geom_point(data = plot_data_points, 
               size = pointsize, 
               aes_string(fill = "method2.satis", colour = "method", shape = "thr"), 
               stroke = 1) + 
    scale_shape_manual(values = rep(21, nthr), guide = "none") + 
    scale_fill_manual(values = newcol, guide = "none", name = "") +
    scale_color_manual(values = newcol,
                       name = "", limits = force) +
    ylim(yaxisrange[1], yaxisrange[2]) + 
    scale_x_continuous(breaks = c(thresholds, seq(0, xaxisrange[2], 0.1)), 
                       labels = c(thresholds, "", seq(0, xaxisrange[2], 0.1)[-1]), 
                       limits = c(xaxisrange[1], xaxisrange[2])) +
    ggtitle(title) +
    theme_grey() + 
    theme(legend.position = "none", 
          panel.background = element_rect(fill = "white", colour = "black"),
          panel.grid.minor.x = element_blank(), 
          panel.grid.minor.y = element_blank(), 
          strip.text = element_text(size = 15), 
          strip.background = element_rect(fill = NA, colour = "black"), 
          axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1, size = 10), 
          axis.text.y = element_text(size = 10), 
          axis.title.x = element_text(size = 12), 
          axis.title.y = element_text(size = 12), 
          plot.title = element_text(face = "bold", colour = "black"))
  
  return(pp)
}
```

# Tasic analysis

## Get metadata

```{r}
# meta-info for counts based analyses
load(file="./Data/Tasic_benchmark_datasets_count.Rdata")
metaInfo_counts <- c(TasicBenchmarkLenient,TasicBenchmarkStringent)
names(metaInfo_counts) <- paste0(names(metaInfo_counts), "_counts")

# meta-info for scaledTPM based analyses
load(file="./Data/Tasic_benchmark_datasets_scaledTPM.Rdata")
metaInfo_sTPM <- c(TasicBenchmarkLenient,TasicBenchmarkStringent)
names(metaInfo_sTPM) <- paste0(names(metaInfo_sTPM), "_sTPM")

metaInfo <- c(metaInfo_counts, metaInfo_sTPM)
```

## Get results

```{r}
# results for counts based analyses for satuRn, DoubleExpSeq and limma 
# diffsplice
load(file="./Data/Tasic_DTU_results_count.Rdata")
TasicDtuBenchmark_counts <- c(TasicDtuBenchmark_satuRn,
                              TasicDtuBenchmark_DoubleExpSeq,
                              TasicDtuBenchmark_limmaDiffsplice)
names(TasicDtuBenchmark_counts) <- paste0(names(TasicDtuBenchmark_counts), 
                                          "_counts")

load(file="./Data/Tasic_DTU_results_scaledTPM.Rdata")
TasicDtuBenchmark_sTPM <- c(TasicDtuBenchmark_satuRn,
                            TasicDtuBenchmark_DoubleExpSeq,
                            TasicDtuBenchmark_limmaDiffsplice)
names(TasicDtuBenchmark_sTPM) <- paste0(names(TasicDtuBenchmark_sTPM), "_sTPM")

results <- c(TasicDtuBenchmark_counts, TasicDtuBenchmark_sTPM)

rm(list=setdiff(ls(), c("results", "metaInfo", "computePlot", "makePlot")))
```

## Generate requested plots

```{r}
plotlist_separate <- computePlot(result = results,
                                 metadata = metaInfo)

count <- 1
figGrobs_Tasic <- list()

for(i in 1:9) {
    plotlist_current <- plotlist_separate[c(count:(count+3))]
    
    gg <- makePlot(plotlist = plotlist_current,
                   colors = c("black", "red", "blue", "green3"),
                   xaxisrange = c(0,0.4),
                   dataset = "Tasic",
                   type = "cells")
    figGrobs_Tasic[[i]] <- gg
    
    count <- count+4
}

saveRDS(figGrobs_Tasic, "./Results/Benchmarks_additional/Filtering_abundanceMetrics/Tasic_strategies_data.rds")
rm(list=setdiff(ls(), c("makePlot", "computePlot")))
```

---

# Darmanis analysis

## Get metadata

```{r}
# meta-info for counts based analyses
load(file="./Data/Darmanis_benchmark_datasets_count.Rdata")
metaInfo_counts <- c(DarmanisBenchmarkLenient,DarmanisBenchmarkStringent)
names(metaInfo_counts) <- paste0(names(metaInfo_counts), "_counts")

# meta-info for scaledTPM based analyses
load(file="./Data/Darmanis_benchmark_datasets_scaledTPM.Rdata")
metaInfo_sTPM <- c(DarmanisBenchmarkLenient,DarmanisBenchmarkStringent)
names(metaInfo_sTPM) <- paste0(names(metaInfo_sTPM), "_sTPM")

metaInfo <- c(metaInfo_counts, metaInfo_sTPM)
```

## Get results

```{r}
# results for counts based analyses for satuRn, DoubleExpSeq and limma diffsplice
load(file="./Data/Darmanis_DTU_results_count.Rdata")
DarmanisDtuBenchmark_counts <- c(DarmanisDtuBenchmark_satuRn,
                                 DarmanisDtuBenchmark_DoubleExpSeq,
                                 DarmanisDtuBenchmark_limmaDiffsplice)
names(DarmanisDtuBenchmark_counts) <- paste0(names(DarmanisDtuBenchmark_counts), "_counts")

load(file="./Data/Darmanis_DTU_results_scaledTPM.Rdata")
DarmanisDtuBenchmark_sTPM <- c(DarmanisDtuBenchmark_satuRn,
                               DarmanisDtuBenchmark_DoubleExpSeq,
                               DarmanisDtuBenchmark_limmaDiffsplice)
names(DarmanisDtuBenchmark_sTPM) <- paste0(names(DarmanisDtuBenchmark_sTPM), "_sTPM")

results <- c(DarmanisDtuBenchmark_counts, DarmanisDtuBenchmark_sTPM)

rm(list=setdiff(ls(), c("results", "metaInfo","makePlot", "computePlot")))
```

## Generate requested plots

```{r}
plotlist_separate <- computePlot(result = results,
                                 metadata = metaInfo)

count <- 1
figGrobs_Darmanis <- list()

for(i in 1:9) {
    plotlist_current <- plotlist_separate[c(count:(count+3))]
    
    gg <- makePlot(plotlist = plotlist_current,
                   colors = c("black", "red", "blue", "green3"),
                   xaxisrange = c(0,0.4),
                   dataset = "Darmanis",
                   type = "cells")
    figGrobs_Darmanis[[i]] <- gg
    
    count <- count+4
}

saveRDS(figGrobs_Darmanis, "./Results/Benchmarks_additional/Filtering_abundanceMetrics/Darmanis_strategies_data.rds")
rm(list=setdiff(ls(), c("makePlot", "computePlot")))
```

---

# Chen analysis

## Get metadata

```{r}
# meta-info for counts based analyses
load(file="./Data/Chen_benchmark_datasets_count.Rdata")
metaInfo_counts <- c(ChenBenchmarkLenient,ChenBenchmarkStringent)
names(metaInfo_counts) <- paste0(names(metaInfo_counts), "_counts")

# meta-info for scaledTPM based analyses
load(file="./Data/Chen_benchmark_datasets_scaledTPM.Rdata")
metaInfo_sTPM <- c(ChenBenchmarkLenient,ChenBenchmarkStringent)
names(metaInfo_sTPM) <- paste0(names(metaInfo_sTPM), "_sTPM")

metaInfo <- c(metaInfo_counts, metaInfo_sTPM)
names(metaInfo)
```

## Get results

```{r}
# results for counts based analyses for satuRn, DoubleExpSeq and limma diffsplice
load(file="./Data/Chen_DTU_results_count.Rdata")
ChenDtuBenchmark_counts <- c(ChenDtuBenchmark_satuRn,
                             ChenDtuBenchmark_DoubleExpSeq,
                             ChenDtuBenchmark_limmaDiffsplice)
names(ChenDtuBenchmark_counts) <- paste0(names(ChenDtuBenchmark_counts), "_counts")

load(file="./Data/Chen_DTU_results_scaledTPM.Rdata")
ChenDtuBenchmark_sTPM <- c(ChenDtuBenchmark_satuRn,
                           ChenDtuBenchmark_DoubleExpSeq,
                           ChenDtuBenchmark_limmaDiffsplice)
names(ChenDtuBenchmark_sTPM) <- paste0(names(ChenDtuBenchmark_sTPM), "_sTPM")

results <- c(ChenDtuBenchmark_counts, ChenDtuBenchmark_sTPM)

rm(list=setdiff(ls(), c("results", "metaInfo","makePlot", "computePlot")))
```

## Generate requested plots

```{r}
plotlist_separate <- computePlot(result = results,
                                 metadata = metaInfo)

count <- 1
figGrobs_Chen <- list()

for(i in 1:6) {
    plotlist_current <- plotlist_separate[c(count:(count+3))]
    
    gg <- makePlot(plotlist = plotlist_current,
                   colors = c("black", "red", "blue", "green3"),
                   xaxisrange = c(0,0.4),
                   dataset = "Chen",
                   type = "cells")
    figGrobs_Chen[[i]] <- gg
    
    count <- count+4
}

saveRDS(figGrobs_Chen, "./Results/Benchmarks_additional/Filtering_abundanceMetrics/Chen_strategies_data.rds")
rm(list=setdiff(ls(), c("makePlot", "computePlot")))
```

---

# Gtex analysis

## Get metadata

```{r}
# meta-info for counts based analyses
load(file="./Data/GTEx_benchmark_datasets_count.Rdata")
metaInfo_counts <- c(gtexBenchmarkDataLenient,gtexBenchmarkDataStringent)
names(metaInfo_counts) <- paste0(names(metaInfo_counts), "_counts")

# meta-info for scaledTPM based analyses
load(file="./Data/GTEx_benchmark_datasets_scaledTPM.Rdata")
metaInfo_sTPM <- c(gtexBenchmarkDataLenient,gtexBenchmarkDataStringent)
names(metaInfo_sTPM) <- paste0(names(metaInfo_sTPM), "_sTPM")

metaInfo <- c(metaInfo_counts, metaInfo_sTPM)
```

## Get results

```{r}
# results for counts based analyses for satuRn, DoubleExpSeq and limma diffsplice
load(file="./Data/GTEx_DTU_results_count.Rdata")
gtexDtuBenchmark_counts <- c(gtexDtuBenchmark_satuRn,
                             gtexDtuBenchmark_DoubleExpSeq,
                             gtexDtuBenchmark_limmaDiffsplice)
names(gtexDtuBenchmark_counts) <- paste0(names(gtexDtuBenchmark_counts), "_counts")

load(file="./Data/GTEx_DTU_results_scaledTPM.Rdata")
gtexDtuBenchmark_sTPM <- c(gtexDtuBenchmark_satuRn,
                           gtexDtuBenchmark_DoubleExpSeq,
                           gtexDtuBenchmark_limmaDiffsplice)
names(gtexDtuBenchmark_sTPM) <- paste0(names(gtexDtuBenchmark_sTPM), "_sTPM")

results <- c(gtexDtuBenchmark_counts, gtexDtuBenchmark_sTPM)

rm(list=setdiff(ls(), c("results", "metaInfo","makePlot", "computePlot")))
```

## Generate requested plots

```{r}
plotlist_separate <- computePlot(result = results,
                                 metadata = metaInfo)

count <- 1
figGrobs_gtex <- list()

for(i in 1:9) {
    plotlist_current <- plotlist_separate[c(count:(count+3))]
    
    gg <- makePlot(plotlist = plotlist_current,
                   colors = c("black", "red", "blue", "green3"),
                   xaxisrange = c(0,0.4),
                   dataset = "Gtex",
                   type = "samples")
    figGrobs_gtex[[i]] <- gg
    
    count <- count+4
}

saveRDS(figGrobs_gtex, "./Results/Benchmarks_additional/Filtering_abundanceMetrics/gtex_strategies_data.rds")
rm(list=setdiff(ls(), c("makePlot", "computePlot")))
```

# Hsapiens analysis

## Get metadata

```{r}
# meta-info for counts based analyses
load(file="./Data/Hsapiens_benchmark_datasets_count.Rdata")
metaInfo_counts <- list(HsBenchmarkLenient,HsBenchmarkStringent)
names(metaInfo_counts) <- c("samples_used_5_rep_1_filterLenient_counts", 
                            "samples_used_5_rep_1_filterStringent_counts")

# meta-info for scaledTPM based analyses
load(file="./Data/Hsapiens_benchmark_datasets_scaledTPM.Rdata")
metaInfo_sTPM <- list(HsBenchmarkLenient,HsBenchmarkStringent)
names(metaInfo_sTPM) <- c("samples_used_5_rep_1_filterLenient_sTPM", 
                          "samples_used_5_rep_1_filterStringent_sTPM")

metaInfo <- c(metaInfo_counts, metaInfo_sTPM)
for(i in seq_along(metaInfo)){
    colnames(metaInfo[[i]]$metaInfo)[3] <- "txSwapped"
}
```

## Get results

```{r}
# results for counts based analyses for satuRn, DoubleExpSeq and limma diffsplice
load(file="./Data/Hsapiens_DTU_results_count.Rdata")
HsapiensDtuBenchmark_counts <- c(HsDtuBenchmark_satuRn,
                                 HsDtuBenchmark_DoubleExpSeq,
                                 HsDtuBenchmark_limmaDiffsplice)

filter <- rep(c("filterLenient","filterStringent"), times=3)
method <- rep(c("satuRn","DoubleExpSeq","limma_diffsplice"), each=2)
names(HsapiensDtuBenchmark_counts) <- paste0(method, 
                                             "_samples_used_5_rep_1_",
                                             filter,
                                             "_counts")

load(file="./Data/Hsapiens_DTU_results_scaledTPM.Rdata")
HsapiensDtuBenchmark_sTPM <- c(HsDtuBenchmark_satuRn,
                               HsDtuBenchmark_DoubleExpSeq,
                               HsDtuBenchmark_limmaDiffsplice)
names(HsapiensDtuBenchmark_sTPM) <- paste0(method,
                                           "_samples_used_5_rep_1_",
                                           filter,
                                           "_sTPM")
results <- c(HsapiensDtuBenchmark_counts, HsapiensDtuBenchmark_sTPM)

rm(list=setdiff(ls(), c("results", "metaInfo","makePlot", "computePlot")))
```

## Generate requested plots

```{r}
plotlist_separate <- computePlot(result = results,
                                 metadata = metaInfo)

count <- 1
figGrobs_Hsapiens <- list()

for(i in 1:3) {
    plotlist_current <- plotlist_separate[c(count:(count+3))]
    
    gg <- makePlot(plotlist = plotlist_current,
                   colors = c("black", "red", "blue", "green3"),
                   xaxisrange = c(0,0.4),
                   dataset = "Hsapiens",
                   type = "samples")
    figGrobs_Hsapiens[[i]] <- gg
    
    count <- count+4
}

saveRDS(figGrobs_Hsapiens, "./Results/Benchmarks_additional/Filtering_abundanceMetrics/Hsapiens_strategies_data.rds")
rm(list=setdiff(ls(), c("makePlot", "computePlot")))
```

# Dmelanogaster analysis

## Get metadata

```{r}
# meta-info for counts based analyses
load(file="./Data/Dmelanogaster_benchmark_datasets_count.Rdata")
metaInfo_counts <- list(DmBenchmarkLenient,DmBenchmarkStringent)
names(metaInfo_counts) <- c("samples_used_5_rep_1_filterLenient_counts", 
                            "samples_used_5_rep_1_filterStringent_counts")

# meta-info for scaledTPM based analyses
load(file="./Data/Dmelanogaster_benchmark_datasets_scaledTPM.Rdata")
metaInfo_sTPM <- list(DmBenchmarkLenient,DmBenchmarkStringent)
names(metaInfo_sTPM) <- c("samples_used_5_rep_1_filterLenient_sTPM", 
                          "samples_used_5_rep_1_filterStringent_sTPM")

metaInfo <- c(metaInfo_counts, metaInfo_sTPM)
for(i in seq_along(metaInfo)){
    colnames(metaInfo[[i]]$metaInfo)[3] <- "txSwapped"
}
```

## Get results

```{r}
# results for counts based analyses for satuRn, DoubleExpSeq and limma diffsplice
load(file="./Data/Dmelanogaster_DTU_results_count.Rdata")
DmelanogasterDtuBenchmark_counts <- c(DmDtuBenchmark_satuRn,
                                      DmDtuBenchmark_DoubleExpSeq,
                                      DmDtuBenchmark_limmaDiffsplice)

filter <- rep(c("filterLenient","filterStringent"), times=3)
method <- rep(c("satuRn","DoubleExpSeq","limma_diffsplice"), each=2)
names(DmelanogasterDtuBenchmark_counts) <- paste0(method, 
                                             "_samples_used_5_rep_1_",
                                             filter,
                                             "_counts")

load(file="./Data/Dmelanogaster_DTU_results_scaledTPM.Rdata")
DmelanogasterDtuBenchmark_sTPM <- c(DmDtuBenchmark_satuRn,
                                      DmDtuBenchmark_DoubleExpSeq,
                                      DmDtuBenchmark_limmaDiffsplice)
names(DmelanogasterDtuBenchmark_sTPM) <- paste0(method,
                                           "_samples_used_5_rep_1_",
                                           filter,
                                           "_sTPM")
results <- c(DmelanogasterDtuBenchmark_counts, DmelanogasterDtuBenchmark_sTPM)

rm(list=setdiff(ls(), c("results", "metaInfo","makePlot", "computePlot")))
```

## Generate requested plots

```{r}
plotlist_separate <- computePlot(result = results,
                                 metadata = metaInfo)

count <- 1
figGrobs_Dmelanogaster <- list()

for(i in 1:3) {
    plotlist_current <- plotlist_separate[c(count:(count+3))]
    
    gg <- makePlot(plotlist = plotlist_current,
                   colors = c("black", "red", "blue", "green3"),
                   xaxisrange = c(0,0.4),
                   dataset = "Dmelanogaster",
                   type = "samples")
    figGrobs_Dmelanogaster[[i]] <- gg
    
    count <- count+4
}

saveRDS(figGrobs_Dmelanogaster, "./Results/Benchmarks_additional/Filtering_abundanceMetrics/Dmelanogaster_strategies_data.rds")
rm(list=setdiff(ls(), c("makePlot", "computePlot")))
```

# Love analysis

## Get metadata

```{r}
# meta-info for counts based analyses
load(file="./Data/Love_benchmark_datasets_count.Rdata")
metaInfo_counts <- c(LoveBenchmarkLenient,LoveBenchmarkStringent)
names(metaInfo_counts) <- paste0(names(metaInfo_counts), "_counts")

# meta-info for scaledTPM based analyses
load(file="./Data/Love_benchmark_datasets_scaledTPM.Rdata")
metaInfo_sTPM <- c(LoveBenchmarkLenient,LoveBenchmarkStringent)
names(metaInfo_sTPM) <- paste0(names(metaInfo_sTPM), "_sTPM")

metaInfo <- c(metaInfo_counts, metaInfo_sTPM)
for(i in seq_along(metaInfo)){
    colnames(metaInfo[[i]]$metaInfo)[3] <- "txSwapped"
}
```

## Get results

```{r}
# results for counts based analyses for satuRn, DoubleExpSeq and limma diffsplice
load(file="./Data/Love_DTU_results_count.Rdata")
LoveDtuBenchmark_counts <- c(loveDtuBenchmark_satuRn,
                             loveDtuBenchmark_DoubleExpSeq,
                             loveDtuBenchmark_limmaDiffsplice)
names(LoveDtuBenchmark_counts) <- paste0(names(LoveDtuBenchmark_counts), "_counts")

load(file="./Data/Love_DTU_results_scaledTPM.Rdata")
LoveDtuBenchmark_sTPM <- c(loveDtuBenchmark_satuRn,
                           loveDtuBenchmark_DoubleExpSeq,
                           loveDtuBenchmark_limmaDiffsplice)
names(LoveDtuBenchmark_sTPM) <- paste0(names(LoveDtuBenchmark_sTPM), "_sTPM")
results <- c(LoveDtuBenchmark_counts, LoveDtuBenchmark_sTPM)

rm(list=setdiff(ls(), c("results", "metaInfo","makePlot", "computePlot")))
```

## Generate requested plots

```{r}
plotlist_separate <- computePlot(result = results,
                                 metadata = metaInfo,
                                 print = FALSE)

count <- 1
figGrobs_Love <- list()

for(i in 1:9) {
    plotlist_current <- plotlist_separate[c(count:(count+3))]
    
    gg <- makePlot(plotlist = plotlist_current,
                   colors = c("black", "red", "blue", "green3"),
                   xaxisrange = c(0,0.4),
                   dataset = "Love",
                   type = "samples")
    figGrobs_Love[[i]] <- gg
    
    count <- count+4
}

saveRDS(figGrobs_Love, "./Results/Benchmarks_additional/Filtering_abundanceMetrics/Love_strategies_data.rds")
rm(list=ls())
gc()
```

# Plot to PDF

```{r}
figGrobs_Tasic <- readRDS("./Results/Benchmarks_additional/Filtering_abundanceMetrics/Tasic_strategies_data.rds")
figGrobs_Darmanis <- readRDS("./Results/Benchmarks_additional/Filtering_abundanceMetrics/Darmanis_strategies_data.rds")
figGrobs_Chen <- readRDS("./Results/Benchmarks_additional/Filtering_abundanceMetrics/Chen_strategies_data.rds")
figGrobs_gtex <- readRDS("./Results/Benchmarks_additional/Filtering_abundanceMetrics/gtex_strategies_data.rds")
figGrobs_Hsapiens <- readRDS("./Results/Benchmarks_additional/Filtering_abundanceMetrics/Hsapiens_strategies_data.rds")
figGrobs_Dmelanogaster <- readRDS("./Results/Benchmarks_additional/Filtering_abundanceMetrics/Dmelanogaster_strategies_data.rds")
figGrobs_Love <- readRDS("./Results/Benchmarks_additional/Filtering_abundanceMetrics/Love_strategies_data.rds")

# here each panel is plotted on a separate PDF page
pdf(file="./Results/Benchmarks_additional/Filtering_abundanceMetrics/filt_quant_all.pdf",
    width = 5,
    height = 5)

figGrobs_Dmelanogaster
figGrobs_Love
figGrobs_Hsapiens
figGrobs_gtex
figGrobs_Tasic
figGrobs_Chen
figGrobs_Darmanis

dev.off()

# alternatively, use grid.arrange to plot multiple panels on 1 page
# grid.arrange(figGrobs_Tasic[[7]], figGrobs_Tasic[[9]],
#              figGrobs_Tasic[[1]], figGrobs_Tasic[[3]], 
#              figGrobs_Tasic[[4]], figGrobs_Tasic[[6]],
#              ncol=2, nrow=3)
```


