---
title: "Additional benchmark: stratify performance on data sparsity"
author: "Jeroen Gilis"
date: "08/07/2022"
output: html_document
---

In this script, we explored how differences in sparsity between the different 
datasets are reflected in the performance benchmarks. To this end, we construct 
FDR-TPR curves where the genes (not individual transcripts) have been stratified
on the percentage of zero counts. For each dataset, we thus assess the 
performance on three strata of genes, i.e., for genes that have a low (< 25%), 
middle (25-50%) or high (> 50%) percentage of zero counts in their corresponding transcript-level count matrices. All the transcripts of the gene get stratified 
in this gene-level stratum. The results of this performance benchmark have been 
included in the Extended data, Figures S30 and S31. As expected, we observe that 
performances are higher when there is a low percentage of zero counts, uniformly 
across all datasets and for all methods. Second, performances on genes with high 
percentage zero counts increase strongly when the number of cells in the data 
increases (comparison between the bottom rows of Figures S30 and S31). Third, 
some methods seem more sensitive than others against sparse data. In particular,
the performance of DEXSeq on the Tasic et al. 20v20 dataset and the Chen et al. 
20v20 dataset (Figure S30) is on par with the performance of satuRn and 
DoubleExpSeq for the strata with a low or middle percentage of zeros, but 
markedly lower for the stratum with the highest percentage of zeros. As such, 
the performance gain of satuRn and DoubleExpSeq over DEXSeq in these datasets 
can be largely attributed to a better performance on the sparsest part of the 
data. Note that this is in line with the performance of DEXSeq being on par 
with satuRn and DoubleExpSeq on bulk data.

```{r setup, include=FALSE, echo=FALSE}
require("knitr")
opts_knit$set(root.dir = rprojroot::find_rstudio_root_file())
```

# Load libraries

```{r,message=FALSE,warning=FALSE}
library(ggplot2)
library(iCOBRA)
```

# Helper functions

## get_pvalues

Get pvalues for the different methods on the different datasets

```{r}
get_pvalues <- function(dataset,nrep,nmethods) {
  
  get_all_txs <- function(dataset,nrep){
  
    all_txs <- c()
    for (i in 1:nrep) {
      current_txs <- as.character(dataset[[i]]$dtuAnalysis$TXNAME)
      all_txs <- c(all_txs, current_txs)
      all_txs <- unique(all_txs)
    }
    return(all_txs)
  }

  all_txs <- get_all_txs(dataset,nrep)
  pvalues <- matrix(data=NA, nrow=length(all_txs), ncol=nmethods*nrep)
  rownames(pvalues) <- all_txs
  
  for (i in 1:(nmethods*nrep)) {
    pvalues_new <- data.frame(dataset[[i]]$dtuAnalysis$p_value)
    rownames(pvalues_new) <- as.character(dataset[[i]]$dtuAnalysis$TXNAME)
    colnames(pvalues_new) <- names(dataset)[i]

    pvalues[,i] <- pvalues_new[match(rownames(pvalues),rownames(pvalues_new)),1]
  }
  
  pvalues <- apply(pvalues, 2, as.numeric)
  rownames(pvalues) <- all_txs
  pvalues <- as.data.frame(pvalues)
  pvalues <- pvalues[order(rownames(pvalues)),]
  
  if(nmethods == 6){
    colnames(pvalues) <- c("satuRn", "limma_diffsplice", "DEXSeq", 
                           "DRIMSeq","edgeR_diffsplice", "DoubleExpSeq")
  } else {
    colnames(pvalues) <- c("satuRn", "limma_diffsplice",
                           "edgeR_diffsplice", "DoubleExpSeq")
  }
  return(pvalues)
}
```

## visualize_datasets

Generate FDR TPR curves with performances for each method stratified on the 
number of zero counts

```{r}
visualize_datasets <- function(dataset,
                               metaInfo,
                               saveName,
                               nFilters,
                               nSampleSizes,
                               nRep,
                               nmethods,
                               selection,
                               selection2) {

    count <- 0
    nPlots <- nFilters * nSampleSizes
    print(paste("This function will generate",  nPlots, "plots"))
  
    for (j in 1:nPlots) {
        print(j)
        Tasic_current_outer <- dataset[grepl(selection[j], names(dataset))]
        metaInfo_current_outer <- metaInfo[grepl(selection2[j], names(metaInfo))]
        sampleSize <- gsub("([0-9]+).*$", "\\1", selection2[j])

        resList <- c()
        count <- count+1

        for (k in 1:length(nRep)) {
            Tasic_current_inner <- Tasic_current_outer[grepl(nRep[k], names(Tasic_current_outer))]

            metaInfo_current_inner <- metaInfo_current_outer[grepl(nRep[k], names(metaInfo_current_outer))]
    
            pvalues <- get_pvalues(Tasic_current_inner,nrep=1,nmethods)

            truth_file <- Tasic_current_inner[[1]]$dtuAnalysis
    
            if(count <= nSampleSizes){
                txSwapped <- metaInfo_current_inner[1]          
                } else {
                txSwapped <- metaInfo_current_inner[2]
            }
    
            truth_file <- truth_file[,c("TXNAME","p_value")]
            
            ################### extra part to compute splv's ###################
            counts <- txSwapped[[1]]$data
            tx2gene <- txSwapped[[1]]$metaInfo[,c(1,2)]
            ncells <- ncol(counts)
            frac0 <- c()
            
            for (gene in unique(tx2gene$GENEID)) {

                transcripts <- tx2gene[tx2gene$GENEID == gene, "TXNAME"]
                frac0 <- c(frac0,rep(
                           sum(counts[transcripts,] == 0)/(length(transcripts)*ncells),times=length(transcripts)))
            }
            
            names(frac0) <- tx2gene$TXNAME
            frac0_cat <- cut(frac0, 
                             breaks=c(-Inf, 0.25, 0.5, Inf), 
                             labels=c("low","middle","high"))
            names(frac0_cat) <- names(frac0)
            ####################################################################
    
            txSwapped[[1]]$metaInfo <- txSwapped[[1]]$metaInfo[match(truth_file$TXNAME,txSwapped[[1]]$metaInfo$TXNAME),]
    
            truth_file$gene_modified <- as.numeric(txSwapped[[1]]$metaInfo$txSwapped)
            truth_file <- truth_file[order(truth_file$TXNAME),]
            truth_file$frac0_cat <- frac0_cat[match(truth_file$TXNAME, names(frac0_cat))]
            truth_file$frac0_cat <- as.factor(truth_file$frac0_cat)
            rownames(truth_file) <- truth_file$TXNAME

            resList[[k]] <- list(pvalues,truth_file)
        }
        
        rownames(resList[[1]][[1]]) <- paste0(rownames(resList[[1]][[1]]), "_1")
        rownames(resList[[2]][[1]]) <- paste0(rownames(resList[[2]][[1]]), "_2")
        rownames(resList[[3]][[1]]) <- paste0(rownames(resList[[3]][[1]]), "_3")
        
        pvalues_full <- rbind(resList[[1]][[1]],
                              resList[[2]][[1]],
                              resList[[3]][[1]])
        
        rownames(resList[[1]][[2]]) <- paste0(rownames(resList[[1]][[2]]), "_1")
        rownames(resList[[2]][[2]]) <- paste0(rownames(resList[[2]][[2]]), "_2")
        rownames(resList[[3]][[2]]) <- paste0(rownames(resList[[3]][[2]]), "_3")
        
        truth_file_full <- rbind(resList[[1]][[2]],
                                 resList[[2]][[2]],
                                 resList[[3]][[2]])
        
        lowlevel <- paste0("low (n=",round(sum(truth_file_full$frac0_cat == "low")/3),")")
        middlelevel <- paste0("middle (n=",round(sum(truth_file_full$frac0_cat == "middle")/3),")")
        highlevel <- paste0("high (n=",round(sum(truth_file_full$frac0_cat == "high")/3),")")
        
        levels(truth_file_full$frac0_cat) <- c(lowlevel,middlelevel,highlevel)
        
        cobra <- COBRAData(pval = pvalues_full, truth = truth_file_full)
        cobra <- calculate_adjp(cobra)
        cobra1perf <- calculate_performance(cobra, binary_truth = "gene_modified", cont_truth = "none", splv = "frac0_cat", aspects = c("fdrtpr", "fdrtprcurve", "overlap"))

        cobraplot <- prepare_data_for_plot(cobra1perf, colorscheme = "Dark2", facetted = TRUE)
        cobraplot@fdrtprcurve$splitval <- factor(cobraplot@fdrtprcurve$splitval,
                                                 levels=c("overall",
                                                          paste0("frac0_cat:",lowlevel),
                                                          paste0("frac0_cat:",middlelevel),
                                                          paste0("frac0_cat:",highlevel)))

        if (nmethods == 6){
          
            colors <- c("#E69F00", "#56B4E9", "#009E73", "#F0E442", "#0072B2", "black", "#999999")
            whites <- length(which(cobraplot@plotcolors == "white"))
            notwhites <- length(which(cobraplot@plotcolors != "white"))/length(colors)
            new_col <- c(rep(colors,notwhites),
                        rep("white",whites))
            names(new_col) <- names(cobraplot@plotcolors)
            cobraplot@plotcolors <- new_col
    
            plot <- plot_fdrtprcurve(cobraplot, 
                                     xaxisrange = c(0, 0.4), 
                                     yaxisrange = c(0,1),
                                     pointsize = 2,
                                     linewidth = 0.75)
            plot <- plot + scale_color_manual(values = new_col, limits=force)
    
            current_title <- paste0(sampleSize, " versus ", 
                                    sampleSize, " - edgeR filter - count")
        } else {
            colors <- c("#56B4E9", "#F0E442", "#0072B2", "black", "#999999")
            whites <- length(which(cobraplot@plotcolors == "white"))
            notwhites <- length(which(cobraplot@plotcolors != "white"))/length(colors)
            
            new_col <-c(rep(colors,notwhites),
                        rep("white", whites))
            names(new_col) <- names(cobraplot@plotcolors)
            cobraplot@plotcolors <- new_col
    
            plot <- plot_fdrtprcurve(cobraplot, 
                                     xaxisrange = c(0, 0.3), 
                                     yaxisrange = c(0,1),
                                     pointsize = 2,
                                     linewidth = 0.75)
            plot <- plot + scale_color_manual(values = new_col, limits=force)
            current_title <- paste0(sampleSize, " versus ", 
                                    sampleSize, " - edgeR filter - count")
        }

        plot <- plot + 
            ggtitle(current_title) +
            theme(strip.text.x = element_text(size = 12),
                  plot.title = element_text(size=14,face="bold"),
                  axis.text.x = element_text(size = 12),
                  axis.text.y = element_text(size = 12),
                  axis.title.x = element_text(size = 12),
                  axis.title.y = element_text(size = 12),
                  legend.position="none") +
            facet_wrap(~splitval, ncol = 1)
      
      png(paste("./Results/Stratify_zeroes/", saveName, "_",  selection2[j], "_",
                sub("^[^f]*", "", selection[j]), "_count.png", sep = ""),
            width     = 3.5,
            height    = 12,
            units     = "in",
            res       = 200,
            pointsize = 4) # start export
      print(plot)
      dev.off()
  }
}
```

# Tasic dataset

## Load truth file

```{r}
load(file="./Data/Tasic_benchmark_datasets_count.Rdata")
metaInfo <- c(TasicBenchmarkLenient,TasicBenchmarkStringent)
rm(TasicBenchmarkLenient,TasicBenchmarkStringent)
```

## Load DTU performance data

```{r}
load(file="./Data/Tasic_DTU_results_count.Rdata")

TasicBenchmark <- c(
    TasicDtuBenchmark_satuRn,
    TasicDtuBenchmark_limmaDiffsplice,
    TasicDtuBenchmark_DEXSeq,
    TasicDtuBenchmark_DRIMSeq,
    TasicDtuBenchmark_edgeRdiffsplice,
    TasicDtuBenchmark_DoubleExpSeq)

### Remove empty enteries (due to not tested - aka to many samples for reasonable runtime)
TasicBenchmark  <- TasicBenchmark[which(
    sapply(TasicBenchmark, function(x) ! is.null(x$dtuAnalysis)))]

rm(TasicDtuBenchmark_satuRn,
    TasicDtuBenchmark_limmaDiffsplice,
    TasicDtuBenchmark_DEXSeq,
    TasicDtuBenchmark_DRIMSeq,
    TasicDtuBenchmark_edgeRdiffsplice,
    TasicDtuBenchmark_DoubleExpSeq)
invisible(gc())
```

## Generate FDR-TPR curves

```{r}
selection <- c("20_rep_._filterLenient")
selection2 <- c("20_rep")

visualize_datasets(dataset=TasicBenchmark,
                   metaInfo=metaInfo,
                   saveName="Tasic",
                   nFilters=1,
                   nSampleSizes=1,
                   nRep=c("rep_1","rep_2","rep_3"),
                   nmethods=6,
                   selection=selection,
                   selection2=selection2)

selection <- c("200_rep_._filterLenient")
selection2 <- c("200_rep")

visualize_datasets(TasicBenchmark,
                   metaInfo,
                   saveName="Tasic",
                   nFilters=1,
                   nSampleSizes=1,
                   nRep=c("rep_1","rep_2","rep_3"),
                   nmethods=4,
                   selection = selection,
                   selection2=selection2)
```

# Chen dataset

## Load truth file

```{r}
load(file="./Data/Chen_benchmark_datasets_count.Rdata")
metaInfo <- c(ChenBenchmarkLenient,ChenBenchmarkStringent)
rm(ChenBenchmarkLenient,ChenBenchmarkStringent)
```

## Load DTU performance results

```{r}
load(file="./Data/Chen_DTU_results_count.Rdata")

ChenBenchmark <- c(
    ChenDtuBenchmark_satuRn,
    ChenDtuBenchmark_limmaDiffsplice,
    ChenDtuBenchmark_DEXSeq,
    ChenDtuBenchmark_DRIMSeq,
    ChenDtuBenchmark_edgeRdiffsplice,
    ChenDtuBenchmark_DoubleExpSeq
)

### Remove empty entries (due to not tested - aka to many samples for reasonable runtime)
ChenBenchmark  <- ChenBenchmark[which(
    sapply(ChenBenchmark, function(x) ! is.null(x$dtuAnalysis)))]

rm(ChenDtuBenchmark_satuRn,
    ChenDtuBenchmark_limmaDiffsplice,
    ChenDtuBenchmark_DEXSeq,
    ChenDtuBenchmark_DRIMSeq,
    ChenDtuBenchmark_edgeRdiffsplice,
    ChenDtuBenchmark_DoubleExpSeq)
invisible(gc())
```

# Generate FDR-TPR curves

```{r}
selection <- c("20_rep_._filterLenient")
selection2 <- c("20_rep")

visualize_datasets(dataset=ChenBenchmark,
                   metaInfo,
                   saveName="Chen",
                   nFilters=1,
                   nSampleSizes=1,
                   nRep=c("rep_1","rep_2","rep_3"),
                   nmethods=6,
                   selection = selection,
                   selection2=selection2)


selection <- c("_50_rep_._filterLenient")
selection2 <- c("50_rep")

visualize_datasets(ChenBenchmark,
                   metaInfo,
                   saveName="Chen",
                   nFilters=1,
                   nSampleSizes=1,
                   nRep=c("rep_1","rep_2","rep_3"),
                   nmethods=4,
                   selection = selection,
                   selection2=selection2)
```

# Load truth file

```{r}
load(file="./Data/Darmanis_benchmark_datasets_count.Rdata")
metaInfo <- c(DarmanisBenchmarkLenient,DarmanisBenchmarkStringent)
rm(DarmanisBenchmarkLenient,DarmanisBenchmarkStringent)
```

# Load DTU performance results

```{r}
load(file="./Data/Darmanis_DTU_results_count.Rdata")

DarmanisBenchmark <- c(
    DarmanisDtuBenchmark_satuRn,
    DarmanisDtuBenchmark_limmaDiffsplice,
    DarmanisDtuBenchmark_DEXSeq,
    DarmanisDtuBenchmark_DRIMSeq,
    DarmanisDtuBenchmark_edgeRdiffsplice,
    DarmanisDtuBenchmark_DoubleExpSeq)

### Remove empty entries (due to not tested - aka to many samples for reasonable runtime)
DarmanisBenchmark  <- DarmanisBenchmark[which(
    sapply(DarmanisBenchmark, function(x) ! is.null(x$dtuAnalysis)))]

rm(DarmanisDtuBenchmark_satuRn,
    DarmanisDtuBenchmark_limmaDiffsplice,
    DarmanisDtuBenchmark_DEXSeq,
    DarmanisDtuBenchmark_DRIMSeq,
    DarmanisDtuBenchmark_edgeRdiffsplice,
    DarmanisDtuBenchmark_DoubleExpSeq)
invisible(gc())
```

# Generate FDR-TPR curves

```{r}
selection <- c("20_rep_._filterLenient")
selection2 <- c("20_rep")

visualize_datasets(DarmanisBenchmark,
                   metaInfo,
                   saveName="Darmanis",
                   nFilters=1,
                   nSampleSizes=1,
                   nRep=c("rep_1","rep_2","rep_3"),
                   nmethods=6,
                   selection = selection,
                   selection2=selection2)

selection <- c("100_rep_._filterLenient")
selection2 <- c("100_rep")

visualize_datasets(DarmanisBenchmark,
                   metaInfo,
                   saveName="Darmanis",
                   nFilters=1,
                   nSampleSizes=1,
                   nRep=c("rep_1","rep_2","rep_3"),
                   nmethods=4,
                   selection=selection,
                   selection2=selection2)
```









