---
title: "Comparison_satuRn_limmaDiffsplice"
output: html_document
---

Load libraries

```{r, message=FALSE, warning=FALSE}
library(SummarizedExperiment)
library(qbDTU)
library(ggplot2)
library(tidyverse)
```

Load data

```{r}
sumExp <- readRDS(file="/Users/jg/Desktop/PhD/DTU_project/Github/dtuPaper/Case_study_L5It/sumExp.Rds")
limmaRes_all <- readRDS(file="/Users/jg/Desktop/PhD/DTU_project/Github/dtuPaper/Case_study_L5It/limmaRes_all.Rds")
load("/Users/jg/Desktop/PhD/DTU_project/Github/dtuPaper/Case_study_L5It/topTreat_DGE.Rda")
```

Get ranks (in terms of significance) for each transcript in both mehtods

```{r}
txInfo <- as.data.frame(rowData(sumExp)[,c(1,2)])
txInfo$rank_limma <- match(txInfo$isoform_id,rownames(limmaRes_all[[6]]))
txInfo$rank_qbDTU <- match(txInfo$isoform_id,rownames(rowData(sumExp)[["fitQBResult_C6"]])[order(rowData(sumExp)[["fitQBResult_C6"]]$empirical_pval)])

txInfo$diffRank <- abs(txInfo$rank_limma - txInfo$rank_qbDTU) # difference in ranks between both methods
```

# limmaDiffsplice more often identifies transcripts of genes with low information content

When we inspected the transcripts that were highly ranked in the top DTU list of limma diffsplice but lowly ranked in our top list, we found that most of these transcripts originate from genes that are lowly expressed;

## Global comparison

Get gene-level counts

```{r}
totalCount <- qbDTU:::getTotalCount(assay(sumExp), txInfo)
mean_tot <- rowMeans(totalCount) # summarize gene-level counts as the mean over all cells
```

Identify DTU transcripts in both methods

```{r}
qbDTU_DTU <- rowData(sumExp)[["fitQBResult_C6"]]$empirical_FDR # DTU in qbDTU
names(qbDTU_DTU) <- rownames(rowData(sumExp)[["fitQBResult_C6"]])

limma_DTU <- limmaRes_all[[6]]$empirical_FDR # DTU in limma
names(limma_DTU) <- rownames(limmaRes_all[[6]])
limma_DTU <- limma_DTU[match(names(qbDTU_DTU),names(limma_DTU))]

both <- qbDTU_DTU<0.05 & limma_DTU<0.05 # DTU in both
qbDTU <- qbDTU_DTU<0.05 & !limma_DTU<0.05 # DTU in qbDTU only
limma <- limma_DTU<0.05 & !qbDTU_DTU<0.05 # DTU in limma only
```

get data in correct format for ggplot

```{R}
DTU_all <- limma
DTU_all[which(DTU_all==TRUE)] <- "limmaDiffsplice"
DTU_all[which(qbDTU==TRUE)] <- "satuRn"
DTU_all[which(both==TRUE)] <- "both"
DTU_all[which(DTU_all==FALSE)] <- "none"
gg_data <- as.data.frame(cbind(names(DTU_all),unname(DTU_all),mean_tot))
colnames(gg_data) <- c("transcript","method","totalCount")
gg_data$method <- as.factor(gg_data$method)
gg_data$totalCount <- as.numeric(gg_data$totalCount)
```

Make plot

```{r}
# png("CaseStudy_results/totalCount.png",
#   width     = 5,
#   height    = 5,
#   units     = "in",
#   res       = 200,
#   pointsize = 4)
ggplot(data = gg_data[-which(gg_data$method == "none"),], aes(x=method,y=totalCount,fill=method)) +
  geom_boxplot() +
  scale_y_log10() +
  theme_classic() +
  scale_fill_manual(values=c("grey84","#0072B2","black"))
# dev.off()
```

## Example

As a consequence, transcripts belonging to lowly expressed genes are correctly considered less informative in satuRn and are thus less likely to be picked up. In Figure 7A, we show that while our method estimates a mean usage of 7% in Tnc cells and 26% in Hsd11b1 Endou cells (indicated by the gold diamond), the transcript is not identified as differentially used, given the low abundance of the corresponding gene and the highly dispersed single-cell level observations.

```{r}
group1 <- rownames(colData(sumExp))[colData(sumExp)$group == "ALM.L5_IT_ALM_Tnc"]
group2 <- rownames(colData(sumExp))[colData(sumExp)$group == "VISp.L5_IT_VISp_Hsd11b1_Endou"]

plot_7A <- qbDTU::plotDTU(object = sumExp, contrast = "C6", groups = list(group1,group2), coefficients = list(c(0,0,0,0,0,0,0,0,1,0,0,0,0,0),c(0,0,0,0,0,0,0,0,0,0,0,0,1,0)), summaryStat = c("model","mean"), transcripts = "ENSMUST00000080335",genes = NULL,top.n = 6)

# get same layout for figure as in paper
transcript <- "ENSMUST00000080335"
padj_limma <- format(limmaRes_all[[6]][transcript,"empirical_FDR"],digits=4)
rank_limma <- txInfo[txInfo$isoform_id == transcript,"rank_limma"]

padj_satuRn <- format(rowData(sumExp)[["fitQBResult_C6"]][transcript,"empirical_FDR"],digits=4)
rank_satuRn <- txInfo[txInfo$isoform_id == transcript,"rank_qbDTU"]

label_facet <- function(txID, padj_limma, rank_limma, padj_satuRn, rank_satuRn) 
{
    lev <- levels(as.factor(txID))
    lab <- paste0("limma empFDR = ", padj_limma, ", rank=", rank_limma, "\n", "satuRn empFDR = ", padj_satuRn, ", rank=", rank_satuRn)
    names(lab) <- lev
    return(lab)
}

# plot figure 7A

# png("CaseStudy_results/figure7A.png",
#   width     = 5,
#   height    = 5,
#   units     = "in",
#   res       = 200,
#   pointsize = 4)


plot_7A[[1]] +
  scale_fill_manual(labels = c("ALM","VISp"), values=c("royalblue4", "firebrick")) +
  scale_x_discrete(labels= c("Tnc","Hsd11b1_Endou")) + 
  facet_wrap(~variable, ncol = 1, labeller = labeller(variable = label_facet(factor(transcript), padj_limma, rank_limma, padj_satuRn, rank_satuRn))) + 
              theme(axis.text.x = element_text(angle = 45, vjust = 1, hjust = 1, size = 8)) + 
              theme(strip.text = element_text(size = 7,face = "bold"))

# dev.off()
```

# satuRN more often identifies small changes in transcript usage that are stable across all cells

Conversely, by looking at the transcripts that were highly ranked in our DTU list but lowly ranked in the top list of limma, we observe that our model is more likely to capture small changes in transcript usage that are stable across all cells and belong to genes that are highly expressed

## Example

An example of such a transcript is shown in Figure 7B.

```{r}
plot_7B <- qbDTU::plotDTU(object = sumExp, contrast = "C6", groups = list(group1,group2), coefficients = list(c(0,0,0,0,0,0,0,0,1,0,0,0,0,0),c(0,0,0,0,0,0,0,0,0,0,0,0,1,0)), summaryStat = "model", transcripts = "ENSMUST00000125287",genes = NULL,top.n = 6)

# get same layout for figure as in paper
transcript <- "ENSMUST00000125287"
padj_limma <- format(limmaRes_all[[6]][transcript,"empirical_FDR"],digits=4)
rank_limma <- txInfo[txInfo$isoform_id == transcript,"rank_limma"]

padj_satuRn <- format(rowData(sumExp)[["fitQBResult_C6"]][transcript,"empirical_FDR"],digits=4)
rank_satuRn <- txInfo[txInfo$isoform_id == transcript,"rank_qbDTU"]

# plot figure 7B

# png("CaseStudy_results/figure7B.png",
#   width     = 5,
#   height    = 5,
#   units     = "in",
#   res       = 200,
#   pointsize = 4)

plot_7B[[1]] +
  scale_fill_manual(labels = c("ALM","VISp"), values=c("royalblue4", "firebrick")) +
  scale_x_discrete(labels= c("Tnc","Hsd11b1_Endou")) + 
  facet_wrap(~variable, ncol = 1, labeller = labeller(variable = label_facet(factor(transcript), padj_limma, rank_limma, padj_satuRn, rank_satuRn))) + 
              theme(axis.text.x = element_text(angle = 45, vjust = 1, hjust = 1, size = 8)) + 
              theme(strip.text = element_text(size = 7,face = "bold"))

# dev.off()
```

# limmaDiffsplice is more influenced by outlying observations that have a low gene-level abundance

DTU claims by limma are driven by differences in raw mean usages of transcripts. 

## Example

In Figure 7C, the raw mean usage of the transcript is 77% in Tnc cells and 45% in Hsd11b1 Endou cells, as indicated by the cyan diamonds. By contrast, the mean usage estimate by [name], which takes into account that the Hsd11b1 Endou cells expressing the transcript at 0% usage have low gene-level count, is 83% for Tnc cells and 75% for Hsd11b1 Endou cells, as indicated by the gold diamonds.

```{r}
plot_7C <- qbDTU::plotDTU(object = sumExp, contrast = "C6", groups = list(group1,group2), coefficients = list(c(0,0,0,0,0,0,0,0,1,0,0,0,0,0),c(0,0,0,0,0,0,0,0,0,0,0,0,1,0)), summaryStat = c("model","mean"), transcripts = "ENSMUST00000106956",genes = NULL,top.n = 6)

# get same layout for figure as in paper
transcript <- "ENSMUST00000106956"
padj_limma <- format(limmaRes_all[[6]][transcript,"empirical_FDR"],digits=4)
rank_limma <- txInfo[txInfo$isoform_id == transcript,"rank_limma"]

padj_satuRn <- format(rowData(sumExp)[["fitQBResult_C6"]][transcript,"empirical_FDR"],digits=4)
rank_satuRn <- txInfo[txInfo$isoform_id == transcript,"rank_qbDTU"]

# plot figure 7C

# png("CaseStudy_results/figure7C.png",
#   width     = 5,
#   height    = 5,
#   units     = "in",
#   res       = 200,
#   pointsize = 4)

plot_7C[[1]] +
  scale_fill_manual(labels = c("ALM","VISp"), values=c("royalblue4", "firebrick")) +
  scale_x_discrete(labels= c("Tnc","Hsd11b1_Endou")) + 
  facet_wrap(~variable, ncol = 1, labeller = labeller(variable = label_facet(factor(transcript), padj_limma, rank_limma, padj_satuRn, rank_satuRn))) + 
              theme(axis.text.x = element_text(angle = 45, vjust = 1, hjust = 1, size = 8)) + 
              theme(strip.text = element_text(size = 7,face = "bold"))

# dev.off()
```













