---
title: "Comparison_satuRn_limmaDiffsplice"
author: "Jeroen Gilis"
date: "11/11/2020"
output: html_document
---

This is scripts performs qualitative comparison between the results of the DTU analysis with satuRn and the DTU analysis with limma diffsplice. In order to run this script, two files should be downloaded from Zenodo and placed in the data folder of this GitHub repository;

1. Tasic_caseStudy_satuRn.Rds (satuRn DTU results)
2. Tasic_caseStudy_limmaDiffsplice.Rds (limma diffsplice DTU results)

Note that the metadata files are identical to the one supplemented by Tasic et al., Supplementary_Table_10_Full_Metadata.xlsx and GSE115746_accession_table.csv, respectively. All figures generated in this script are available from our GitHub repository under ./Results/CaseStudy/

```{r setup, include=FALSE, echo=FALSE}
require("knitr")
opts_knit$set(root.dir = rprojroot::find_rstudio_root_file())
```

# Load libraries

```{r, message=FALSE, warning=FALSE}
library(SummarizedExperiment)
library(satuRn)
library(ggplot2)
#library(tidyverse)
```

# Load data

These are the results of the analyses of the Tasic case study with satuRn (generated in satuRn_analysis.Rmd) and with limma diffsplice (generated in limmaDiffsplice_analysis.Rmd).

```{r}
satuRn_results <- readRDS(file="./Data/Tasic_caseStudy_satuRn.Rds")
limmaRes_results <- readRDS(file="./Data/Tasic_caseStudy_limmaDiffsplice.Rds")
```

Get ranks (in terms of significance) for each transcript in both mehtods

```{r}
txInfo <- as.data.frame(rowData(satuRn_results)[,c(1,2)])
txInfo$rank_limma <- match(txInfo$isoform_id,rownames(limmaRes_results[[6]]))
txInfo$rank_satuRn <- match(txInfo$isoform_id,rownames(rowData(satuRn_results)[["fitDTUResult_C6"]])[order(rowData(satuRn_results)[["fitDTUResult_C6"]]$empirical_pval)])

txInfo$diffRank <- abs(txInfo$rank_limma - txInfo$rank_satuRn) # difference in ranks between both methods
```

# limmaDiffsplice more often identifies transcripts of genes with low information content

When we inspected the transcripts that were highly ranked in the top DTU list of limma diffsplice but lowly ranked in our top list, we found that most of these transcripts originate from genes that are lowly expressed;

## Global comparison (for contrast 6)

Get gene-level counts

```{r}
Tasic_counts_current <- assay(satuRn_results)[,satuRn_results$group %in% c("VISp.L5_IT_VISp_Hsd11b1_Endou", "ALM.L5_IT_ALM_Tnc")] # get only cells from contrast 6

totalCount <- satuRn:::getTotalCount(Tasic_counts_current, txInfo)
mean_tot <- rowMeans(totalCount) # summarize gene-level counts as the mean over all cells
```

Get fraction of zeros

```{r}
frac_zero <- rowSums(Tasic_counts_current == 0)/ncol(Tasic_counts_current)
head(frac_zero)
```

Identify DTU transcripts in both methods

```{r}
satuRn_DTU <- rowData(satuRn_results)[["fitDTUResult_C6"]]$empirical_FDR # DTU in satuRn
names(satuRn_DTU) <- rownames(rowData(satuRn_results)[["fitDTUResult_C6"]])

limma_DTU <- limmaRes_results[[6]]$empirical_FDR # DTU in limma
names(limma_DTU) <- rownames(limmaRes_results[[6]])
limma_DTU <- limma_DTU[match(names(satuRn_DTU),names(limma_DTU))]

both <- satuRn_DTU<0.05 & limma_DTU<0.05 # DTU in both
satuRn <- satuRn_DTU<0.05 & !limma_DTU<0.05 # DTU in satuRn only
limma <- limma_DTU<0.05 & !satuRn_DTU<0.05 # DTU in limma only
```

Get data in correct format for ggplot

```{R}
DTU_all <- limma
DTU_all[which(DTU_all==TRUE)] <- "limmaDiffsplice"
DTU_all[which(satuRn==TRUE)] <- "satuRn"
DTU_all[which(both==TRUE)] <- "both"
DTU_all[which(DTU_all==FALSE)] <- "none"
gg_data <- as.data.frame(cbind(names(DTU_all),unname(DTU_all),mean_tot,frac_zero))
colnames(gg_data) <- c("transcript","method","totalCount","frac_zero")
gg_data$method <- as.factor(gg_data$method)
gg_data$totalCount <- as.numeric(as.character(gg_data$totalCount))
gg_data$frac_zero <- as.numeric(as.character(gg_data$frac_zero))

gg_data$method <- factor(gg_data$method, levels = c("satuRn","both","limmaDiffsplice","none"))
```


Make plot of gene-level count of the transcripts uniquely identified as DTU by satuRn or limma diffsplice, or by both methods.

```{r}
png("./Results/caseStudy/Figure7A.png",
  width     = 4.5,
  height    = 3.5,
  units     = "in",
  res       = 400,
  pointsize = 4)
ggplot(data = gg_data[-which(gg_data$method == "none"),], aes(x=method,y=totalCount,fill=method)) +
  geom_boxplot() +
  scale_y_log10() +
  theme_classic() +
  scale_fill_manual(values=c("black","grey84","#0072B2")) +
  ylab("gene-level count") +
  ggtitle("Average gene-level counts of DTU transcripts") +
  theme(plot.title = element_text(size = 10))
dev.off()
```

Make plot of the fraction of non-zero counts in the transcripts uniquely identified as DTU by satuRn or limma diffsplice, or by both methods.

```{r}
gg_data$frac_non_zero <- 1-gg_data$frac_zero

png("./Results/caseStudy/Figure7B.png",
  width     = 4.5,
  height    = 3.5,
  units     = "in",
  res       = 400,
  pointsize = 4)
ggplot(data = gg_data[-which(gg_data$method == "none"),], aes(x=method,y=frac_non_zero,fill=method)) +
  geom_violin() +
  geom_jitter(aes(colour=method),width = 0.1) +
  theme_classic() +
  scale_fill_manual(values=c("black","grey84","#0072B2")) +
  scale_color_manual(values=c("grey60","black","black")) +
  ylim(0,1) +
  ggtitle("Fraction of non-zero counts in DTU transcripts") +
  stat_summary(fun=median, geom="point", shape=18, size=5, col="darkgreen") +
  ylab("fraction_non_zero")  +
  theme(plot.title = element_text(size = 10))
dev.off()
```

## Examples

As a consequence, transcripts belonging to lowly expressed genes are correctly considered less informative in satuRn and are thus less likely to be picked up. In Figure 8A, we show that while our method estimates a mean usage of 7% in Tnc cells and 26% in Hsd11b1 Endou cells (indicated by the gold diamond), the transcript is not identified as differentially used, given the low abundance of the corresponding gene and the highly dispersed single-cell level observations.

```{r}
group1 <- rownames(colData(satuRn_results))[colData(satuRn_results)$group == "ALM.L5_IT_ALM_Tnc"]
group2 <- rownames(colData(satuRn_results))[colData(satuRn_results)$group == "VISp.L5_IT_VISp_Hsd11b1_Endou"]

plot_8A <- satuRn::plotDTU(object = satuRn_results, contrast = "C6", groups = list(group1,group2), coefficients = list(c(0,0,0,0,0,0,0,0,1,0,0,0,0,0),c(0,0,0,0,0,0,0,0,0,0,0,0,1,0)), summaryStat = c("model","mean"), transcripts = "ENSMUST00000080335",genes = NULL,top.n = 6)

# get same layout for figure as in paper
transcript <- "ENSMUST00000080335"
padj_limma <- format(limmaRes_results[[6]][transcript,"empirical_FDR"],digits=4)
rank_limma <- txInfo[txInfo$isoform_id == transcript,"rank_limma"]

padj_satuRn <- format(rowData(satuRn_results)[["fitDTUResult_C6"]][transcript,"empirical_FDR"],digits=4)
rank_satuRn <- txInfo[txInfo$isoform_id == transcript,"rank_satuRn"]

label_facet <- function(txID, padj_limma, rank_limma, padj_satuRn, rank_satuRn) 
{
    lev <- levels(as.factor(txID))
    lab <- paste0("limma empFDR = ", padj_limma, ", rank=", rank_limma, "\n", "satuRn empFDR = ", padj_satuRn, ", rank=", rank_satuRn)
    names(lab) <- lev
    return(lab)
}

# plot figure 8A
png("./Results/caseStudy/Figure8A.png",
  width     = 5,
  height    = 5,
  units     = "in",
  res       = 200,
  pointsize = 4)
plot_8A[[1]] +
  scale_fill_manual(labels = c("ALM","VISp"), values=c("royalblue4", "firebrick")) +
  scale_x_discrete(labels= c("Tnc","Hsd11b1_Endou")) + 
  facet_wrap(~variable, ncol = 1, labeller = labeller(variable = label_facet(factor(transcript), padj_limma, rank_limma, padj_satuRn, rank_satuRn))) + 
              theme(axis.text.x = element_text(angle = 0, vjust = 1, hjust = 0.5, size = 9)) + 
              theme(strip.text = element_text(size = 8,face = "bold"))

dev.off()
```

# satuRN more often identifies small changes in transcript usage that are stable across all cells

Conversely, by looking at the transcripts that were highly ranked in our DTU list but lowly ranked in the top list of limma, we observe that our model is more likely to capture small changes in transcript usage that are stable across all cells and belong to genes that are highly expressed

## Example

An example of such a transcript is shown in Figure 8B.

```{r}
plot_8B <- satuRn::plotDTU(object = satuRn_results, contrast = "C6", groups = list(group1,group2), coefficients = list(c(0,0,0,0,0,0,0,0,1,0,0,0,0,0),c(0,0,0,0,0,0,0,0,0,0,0,0,1,0)), summaryStat = "model", transcripts = "ENSMUST00000125287",genes = NULL,top.n = 6)

# get same layout for figure as in paper
transcript <- "ENSMUST00000125287"
padj_limma <- format(limmaRes_results[[6]][transcript,"empirical_FDR"],digits=4)
rank_limma <- txInfo[txInfo$isoform_id == transcript,"rank_limma"]

padj_satuRn <- format(rowData(satuRn_results)[["fitDTUResult_C6"]][transcript,"empirical_FDR"],digits=4)
rank_satuRn <- txInfo[txInfo$isoform_id == transcript,"rank_satuRn"]

# plot figure 8B
png("./Results/caseStudy/Figure8B.png",
  width     = 5,
  height    = 5,
  units     = "in",
  res       = 200,
  pointsize = 4)
plot_8B[[1]] +
  scale_fill_manual(labels = c("ALM","VISp"), values=c("royalblue4", "firebrick")) +
  scale_x_discrete(labels= c("Tnc","Hsd11b1_Endou")) + 
  facet_wrap(~variable, ncol = 1, labeller = labeller(variable = label_facet(factor(transcript), padj_limma, rank_limma, padj_satuRn, rank_satuRn))) + 
              theme(axis.text.x = element_text(angle = 0, vjust = 1, hjust = 0.5, size = 9)) + 
              theme(strip.text = element_text(size = 8,face = "bold"))

dev.off()
```

# limmaDiffsplice is more influenced by outlying observations that have a low gene-level abundance

DTU claims by limma are driven by differences in raw mean usages of transcripts. 

## Example

In Figure 8C, the raw mean usage of the transcript is 77% in Tnc cells and 45% in Hsd11b1 Endou cells, as indicated by the cyan diamonds. By contrast, the mean usage estimate by satuRn, which takes into account that the Hsd11b1 Endou cells expressing the transcript at 0% usage have low gene-level count, is 83% for Tnc cells and 75% for Hsd11b1 Endou cells, as indicated by the gold diamonds.

```{r}
plot_8C <- satuRn::plotDTU(object = satuRn_results, contrast = "C6", groups = list(group1,group2), coefficients = list(c(0,0,0,0,0,0,0,0,1,0,0,0,0,0),c(0,0,0,0,0,0,0,0,0,0,0,0,1,0)), summaryStat = c("model","mean"), transcripts = "ENSMUST00000106956",genes = NULL,top.n = 6)

# get same layout for figure as in paper
transcript <- "ENSMUST00000106956"
padj_limma <- format(limmaRes_results[[6]][transcript,"empirical_FDR"],digits=4)
rank_limma <- txInfo[txInfo$isoform_id == transcript,"rank_limma"]

padj_satuRn <- format(rowData(satuRn_results)[["fitDTUResult_C6"]][transcript,"empirical_FDR"],digits=4)
rank_satuRn <- txInfo[txInfo$isoform_id == transcript,"rank_satuRn"]

# plot figure 8C
png("./Results/caseStudy/Figure8C.png",
  width     = 5,
  height    = 5,
  units     = "in",
  res       = 200,
  pointsize = 4)
plot_8C[[1]] +
  scale_fill_manual(labels = c("ALM","VISp"), values=c("royalblue4", "firebrick")) +
  scale_x_discrete(labels= c("Tnc","Hsd11b1_Endou")) + 
  facet_wrap(~variable, ncol = 1, labeller = labeller(variable = label_facet(factor(transcript), padj_limma, rank_limma, padj_satuRn, rank_satuRn))) + 
              theme(axis.text.x = element_text(angle = 0, vjust = 1, hjust = 0.5, size = 9)) + 
              theme(strip.text = element_text(size = 8,face = "bold"))
dev.off()
```





