---
title: "Comparison_satuRn_DoubleExpSeq"
author: "Jeroen Gilis"
date: "11/11/2020"
output: html_document
---

This is scripts performs qualitative comparison between the results of the DTU analysis with satuRn and the DTU analysis with DoubleExpSeq. In order to run this script, two files should be downloaded from Zenodo and placed in the data folder of this GitHub repository;

1. Tasic_caseStudy_satuRn.Rds (satuRn DTU results)
2. Tasic_caseStudy_DoubleExpSeq.Rds (DoubleExpSeq DTU results)

All figures generated in this script are available from our GitHub repository under ./Results/CaseStudy/

```{r setup, include=FALSE, echo=FALSE}
require("knitr")
opts_knit$set(root.dir = rprojroot::find_rstudio_root_file())
```

# Load libraries

```{r, message=FALSE, warning=FALSE}
library(SummarizedExperiment)
library(satuRn)
library(ggplot2)
#library(tidyverse)
library(locfdr)
```

# Load data

These are the results of the analyses of the Tasic case study with satuRn (generated in satuRn_analysis.Rmd) and DoubleExpSeq (generated in DoubleExpSeq_analysis.Rmd).

```{r}
satuRn_results <- readRDS(file="./Data/Tasic_caseStudy_satuRn.Rds")
DoubleExpSeq_results <- readRDS(file="./Data/Tasic_caseStudy_DoubleExpSeq.Rds")
```

# DoubleExpSeq p-value distribution figure S14

DoubleExpSeq displays a pathological distribution of p-values.

```{r}
for (i in seq_along(DoubleExpSeq_results)) {
    png(paste("./Results/caseStudy/FigureS14_", i, ".png", sep = ""),
            width     = 3.5,
            height    = 2.5,
            units     = "in",
            res       = 300,
            pointsize = 6) # start export
    hist(DoubleExpSeq_results[[i]]$All[,"pVal"],breaks=40,main = paste("contrast", i),xlab="p-values")
    dev.off()
}
```

# DES finds a large number of DTU transcripts

```{r}
for (i in seq_along(DoubleExpSeq_results)) {
    print(sum(DoubleExpSeq_results[[i]]$All[,"Adj.pVal"] < 0.05,na.rm = T))
}
```

# Using locFDR on the results by DoubleExpSeq is problematic

```{r}
DoubleExpSeq_results_locfdr <- list()
for (i in seq_along(DoubleExpSeq_results)) {
  
  direction <- ifelse(DoubleExpSeq_results[[i]]$All[,1]-DoubleExpSeq_results[[i]]$All[,2] <= 0, 1,-1)
  pval <- DoubleExpSeq_results[[i]]$All[,"pVal"]
  FDR <- DoubleExpSeq_results[[i]]$All[,"Adj.pVal"]

  z <- qnorm(pval/2) * sign(direction)
  z_working <- z[!is.na(z)]
  z_working <- z_working[abs(z_working) < 10]
  current_plot <- locfdr(z_working)
  
  ################# start code from locfdr
  N <- length(z_working)
  b <- 4.3 * exp(-0.26 * log(N, 10))
  med <- median(z_working)
  sc <- diff(quantile(z_working)[c(2, 4)])/(2 * qnorm(0.75))
  mlests <- locfdr:::locmle(z_working, xlim = c(med, b * 
      sc))
  lo <- min(z_working)
  up <- max(z_working)
  bre = 120
  breaks <- seq(lo, up, length = bre)
  zzz <- pmax(pmin(z_working, up), lo)
  zh <- hist(zzz, breaks = breaks, plot = F)
  x <- (breaks[-1] + breaks[-length(breaks)])/2
  sw <- 0
  X <- cbind(1, poly(x, df = 7))
  zh <- hist(zzz, breaks = breaks, plot = F)
  y <- zh$counts
  f <- glm(y ~ poly(x, df = 7), poisson)$fit
  Cov.in = list(x = x, X = X, f = f, sw = sw)
  ml.out = locfdr:::locmle(z_working, xlim = c(mlests[1], 
        b * mlests[2]), d = mlests[1], s = mlests[2], Cov.in = Cov.in)
  mlests = ml.out$mle
  ################# end code from locfdr
  
  #mlests <- satuRn:::p.adjust_empirical(z_working,direction)
 
  zval_empirical <- (z - mlests[1])/mlests[2]
  pval_empirical <- 2 * pnorm(-abs(zval_empirical), mean = 0, sd = 1)
  pval_empirical[zval_empirical == -Inf] <- 1
  FDR_empirical <- p.adjust(pval_empirical, method = "BH")
  
  # set NAs
  NAs <- rep(NA, length(names(pval)[-which(names(pval)%in%names(pval_empirical))]))
  names(NAs) <- names(pval)[-which(names(pval)%in%names(pval_empirical))]
  
  pval_empirical <- c(pval_empirical,NAs)
  FDR_empirical <- c(FDR_empirical,NAs)
  
  pval_empirical <- pval_empirical[match(names(pval),names(pval_empirical))]
  FDR_empirical <- FDR_empirical[match(names(pval),names(FDR_empirical))]
  
  DES_empirical <- data.frame(pval,FDR,pval_empirical,FDR_empirical)
  rownames(DES_empirical) <- rownames(DoubleExpSeq_results[[i]]$All)

  DoubleExpSeq_results_locfdr[[i]] <- DES_empirical
}
```

To generate Figure S15

```{r}
direction <- ifelse(DoubleExpSeq_results[[6]]$All[,1]-DoubleExpSeq_results[[6]]$All[,2] <= 0, 1,-1)
pval <- DoubleExpSeq_results[[6]]$All[,"pVal"]
FDR <- DoubleExpSeq_results[[6]]$All[,"Adj.pVal"]

z <- qnorm(pval/2) * sign(direction)
z_working <- z[!is.na(z)]
z_working <- z_working[abs(z_working) < 10]

png("./Results/CaseStudy/FigureS15.png",
        width     = 4.5,
        height    = 3,
        units     = "in",
        res       = 500,
        pointsize = 8)
current_plot <- suppressWarnings(locfdr(z_working))
dev.off()
```

The large number of p-values that are exatly 1 in the DoubleExpSeq results jeopardize using an empirical null to control the number of false positive discoveries.






