---
title: "limmaDiffsplice_analysis"
author: "Jeroen Gilis"
date: "11/11/2020"
output: html_document
---

This is scripts performs a DTU analysis with limma diffsplice on a subset of cells from the single-cell RNA-Seq dataset by Tasic et al. [Tasic paper](https://doi.org/10.1038/s41586-018-0654-5). In order to run this script, three files should be downloaded from Zenodo and placed in the data folder of this GitHub repository;

1. Tasic_caseStudy_transcript_counts.Rds (transcript-level expression matrix)
2. Tasic_metadata_1.xlsx
3. Tasic_metadata_2.csv

Note that the metadata files are identical to the one supplemented by Tasic et al., Supplementary_Table_10_Full_Metadata.xlsx and GSE115746_accession_table.csv, respectively. All figures generated in this script are available from our GitHub repository under ./Results/CaseStudy/

```{r setup, include=FALSE, echo=FALSE}
require("knitr")
opts_knit$set(root.dir = rprojroot::find_rstudio_root_file())
```

# Load libraries

```{r,message=FALSE}
# for data import
library(AnnotationHub)
library(ensembldb)

# for analysis
library(edgeR)
#library(tidyverse)

# for visualization
library(ggplot2)
```

# Data import

## Import transcript information

```{r, message=FALSE}
ah <- AnnotationHub() # Load the annotation resource.
all <- query(ah, "EnsDb") # Query for all available EnsDb databases
ahEdb <- all[["AH75036"]] # for Mus musculus

txs <- transcripts(ahEdb)
tx2gene <- as.data.frame(matrix(data = NA, nrow = length(txs), ncol = 2))
colnames(tx2gene) <- c("TXNAME","GENEID")
tx2gene$TXNAME <- txs$tx_id
tx2gene$GENEID <- txs$gene_id
```

## Import counts

```{r}
Tasic_counts <- readRDS("./Data/Tasic_caseStudy_transcript_counts.Rds")
# Already remove genes with zero counts in all cells
Tasic_counts <- Tasic_counts[rowSums(Tasic_counts) != 0,]
dim(Tasic_counts)
```

## Import metadata

```{r}
access <- read.csv2("./Data/Tasic_metadata_2.csv",sep = "\t")
access <- access[match(colnames(Tasic_counts),access$SRA_Run),]

metaData <- openxlsx::read.xlsx("./Data/Tasic_metadata_1.xlsx")
metaData <- metaData[match(access$sample_name,metaData$sample_name),]

colnames(Tasic_counts) <- metaData$sample_name
metaData <- metaData[metaData$core_intermediate_call == "Core",] # only retain cell that were unambiguously assigned to a certain cell type (cluster)
Tasic_counts <- Tasic_counts[,metaData$sample_name]
metaData <- metaData[,c("sample_name", "brain_region","cluster")]
```

# Data wrangling

```{r}
# Remove transcripts that are the only isoform expressed of a certain gene
txInfo <- tx2gene
colnames(txInfo) <- c('isoform_id','gene_id')
rownames(txInfo) <- NULL

rownames(Tasic_counts) <- sub("\\..*", "", rownames(Tasic_counts))
txInfo <- txInfo[txInfo$isoform_id %in% rownames(Tasic_counts),]
txInfo <- subset(txInfo,duplicated(gene_id) | duplicated(gene_id, fromLast=TRUE))

Tasic_counts <- Tasic_counts[which(rownames(Tasic_counts) %in% txInfo$isoform_id),]
```

# Set up the experimental design

```{r}
metaData$cluster <- gsub(" ", "_", metaData$cluster)
metaData$group <- paste(metaData$brain_region,metaData$cluster,sep=".")

# Remove groups (cell types) with less than 30 cells --> removes 4 groups
remove <- names(table(interaction(metaData$brain_region,metaData$cluster))[table(interaction(metaData$brain_region,metaData$cluster)) < 30])

metaData <- metaData[-which(metaData$group %in% remove),]
group <- as.factor(metaData$group)
 
design <- model.matrix(~ 0 + group) # Factorial design
colnames(design) <- levels(group)

Tasic_counts <- Tasic_counts[,which(colnames(Tasic_counts) %in% metaData$sample_name)]
```

# Filtering

```{r}
filter_all_edgeR_stringent <- filterByExpr(Tasic_counts,
             design = NULL,
             group = metaData$brain_region,
             lib.size = NULL,
             min.count = 10,
             min.total.count = 0,
             large.n = 0,
             min.prop = 0.7)

table(filter_all_edgeR_stringent)
Tasic_counts <- Tasic_counts[filter_all_edgeR_stringent,]
```

```{r}
# Update txInfo according to the filtering procedure
txInfo <- txInfo[which(txInfo$isoform_id %in% rownames(Tasic_counts)),]

# remove transcripts that are the only isoform expressed of a certain gene (after filtering)
txInfo <- subset(txInfo,duplicated(gene_id) | duplicated(gene_id, fromLast=TRUE))

Tasic_counts <- Tasic_counts[which(rownames(Tasic_counts) %in% txInfo$isoform_id),]

# satuRn require the transcripts in the rowData and the transcripts in the count matrix to be in the same order. If not, the resulting models will be matched to the wrong rowData
txInfo <- txInfo[match(rownames(Tasic_counts),txInfo$isoform_id),]

rm(list = setdiff(ls(), c("Tasic_counts", "txInfo", "design", "metaData", "tx2gene","group")))
invisible(gc())
```

# Analysis

## Fit, set contrasts, test

```{r}
geneForEachTx <- txInfo$gene_id[match(rownames(Tasic_counts),txInfo$isoform_id)]
y <- DGEList(counts = Tasic_counts, group=as.factor(metaData$group), genes = geneForEachTx)

y <- calcNormFactors(y)

v <- voom(y, design, plot=F)
fit <- lmFit(v, design)

contrast.matrix <- makeContrasts(VISp.L5_IT_VISp_Batf3-ALM.L5_IT_ALM_Cpa6_Gpr88, VISp.L5_IT_VISp_Col27a1-ALM.L5_IT_ALM_Cbln4_Fezf2, VISp.L5_IT_VISp_Col6a1_Fezf2-ALM.L5_IT_ALM_Cpa6_Gpr88, VISp.L5_IT_VISp_Col6a1_Fezf2-ALM.L5_IT_ALM_Gkn1_Pcdh19, VISp.L5_IT_VISp_Hsd11b1_Endou-ALM.L5_IT_ALM_Lypd1_Gpr88, VISp.L5_IT_VISp_Hsd11b1_Endou-ALM.L5_IT_ALM_Tnc, VISp.L5_IT_VISp_Hsd11b1_Endou-ALM.L5_IT_ALM_Tmem163_Dmrtb1, VISp.L5_IT_VISp_Whrn_Tox2-ALM.L5_IT_ALM_Tmem163_Arhgap25, levels=design)

fit2 <- contrasts.fit(fit, contrast.matrix)
fit2 <- eBayes(fit2)

ex <- limma::diffSplice(fit2, geneid = "genes", verbose = F)

for (i in seq_len(ncol(ex$coefficients))) {
  limmaRes <- topSplice(ex, coef=i, test="t", number = Inf)
  print(sum(limmaRes$FDR <= 0.05))
}
```

## post-processing

Without correction, we obtain a huge number of DTU transcripts with the limma analysis. Therefore, we additionally post-process the results according to the empirical FDR strategy implemented in satuRn (which is adapted from the locfdr function from the locfdr package by Efron et al.).

At the same time, all panels of Figures S12 and S13 of our paper are generated.

```{r}
limmaRes_all<- list()
for (i in seq_len(ncol(ex$coefficients))) {
  limmaRes <- topSplice(ex, coef=i, test="t", number = Inf)
  
  t <- limmaRes$t
  pval <- limmaRes$P.Value
  FDR <- limmaRes$FDR
  
  # Generate Figure S12
  png(paste("./Results/caseStudy/FigureS12_", i, ".png", sep = ""),
          width     = 3.5,
          height    = 2.5,
          units     = "in",
          res       = 300,
          pointsize = 6) # start export
  hist(pval, breaks=40, main = paste("contrast",i), col="white")
  dev.off()
  
  # Generate Figure S13
  zval <- qnorm(pval/2)*sign(t)
  zval_mid <- zval[abs(zval) < 10]
  zval_mid <- zval_mid[!is.na(zval_mid)]
  
  png(paste("./Results/caseStudy/FigureS13_", i, ".png", sep = ""),
          width     = 3.5,
          height    = 2.5,
          units     = "in",
          res       = 300,
          pointsize = 6) # start export
  plot_lfdr <- locfdr::locfdr(zval_mid,plot=T,main=paste("contrast",i))
  dev.off()
  
  empirical <- qbDTU:::p.adjust_empirical(pval,t,plot=F) # empirical FDR correction as implemented in satuRn

  empirical_pval <- empirical$pval
  empirical_FDR <-  empirical$FDR

  result_limma <- data.frame(t,pval,FDR,empirical_pval,empirical_FDR)
  rownames(result_limma) <- rownames(limmaRes)
  
  # print number of DTU transcript after working under the empirical null
  print(sum(result_limma$empirical_FDR <= 0.05))
  
  limmaRes_all[[i]] <- result_limma
}
```

# Save final object

Save resulting object for later comparison of results with satuRn and DoubleExpSeq.

```{r}
saveRDS(limmaRes_all,file = "./Data/Tasic_caseStudy_limmaDiffsplice.Rds")
```

