---
title: "Venn_diagram"
author: "Jeroen Gilis"
date: "04/12/2020"
output: html_document
---

This is scripts compares the DTU results between satuRn, limma diffsplice and DoubleExpSeq by means of a Venn daigram (Figure 9 from the publication). In order to run this script, three files should be downloaded from Zenodo and placed in the data folder of this GitHub repository;

1. Tasic_caseStudy_satuRn.Rds (satuRn DTU results)
2. Tasic_caseStudy_limmaDiffsplice.Rds (limma diffsplice DTU results)
3. Tasic_caseStudy_DoubleExpSeq.Rds (DoubleExpSeq DTU results)

Note that the metadata files are identical to the one supplemented by Tasic et al., Supplementary_Table_10_Full_Metadata.xlsx and GSE115746_accession_table.csv, respectively. All figures generated in this script are available from our GitHub repository under ./Results/CaseStudy/

```{r setup, include=FALSE, echo=FALSE}
require("knitr")
opts_knit$set(root.dir = rprojroot::find_rstudio_root_file())
```

# Load libraries

```{r, message=FALSE, warning=FALSE}
library(SummarizedExperiment)
library(ggplot2)
#library(tidyverse)
library(VennDiagram)
library(grid)
```

# Load data

These are the results of the analyses of the Tasic case study with satuRn (generated in satuRn_analysis.Rmd), limma diffsplice (generated in limmaDiffsplice_analysis.Rmd) and DoubleExpSeq (generated in DoubleExpSeq_analysis.Rmd).

```{r}
satuRn_results <- readRDS(file="./Data/Tasic_caseStudy_satuRn.Rds")
limmaRes_results <- readRDS(file="./Data/Tasic_caseStudy_limmaDiffsplice.Rds")
DoubleExpSeq_results <- readRDS(file="./Data/Tasic_caseStudy_DoubleExpSeq.Rds")
```

# Venn diagram for top200 transcipts in contrast 6

## Extract top200 transcripts in contrast 6

```{r}
# satuRn
satuRn_results_comp6_top200 <- rowData(satuRn_results)[["fitDTUResult_C6"]]
satuRn_results_comp6_top200 <- satuRn_results_comp6_top200[order(satuRn_results_comp6_top200$empirical_pval),][1:200,]
satuRn_results_comp6_top200 <- rownames(satuRn_results_comp6_top200)

# limma diffsplice
limmaRes_results_comp6_top200 <- limmaRes_results[[6]]
limmaRes_results_comp6_top200 <- limmaRes_results_comp6_top200[order(limmaRes_results_comp6_top200$empirical_pval),][1:200,]
limmaRes_results_comp6_top200 <- rownames(limmaRes_results_comp6_top200)

# DoubleExpSeq
DoubleExpSeq_results_comp6_top200 <- as.data.frame(DoubleExpSeq_results[[6]]$All)
DoubleExpSeq_results_comp6_top200 <- DoubleExpSeq_results_comp6_top200[order(DoubleExpSeq_results_comp6_top200$pVal),][1:200,]
DoubleExpSeq_results_comp6_top200 <- rownames(DoubleExpSeq_results_comp6_top200)
```

## create Venn diagram

Venn diagrams

```{r}
png("./Results/CaseStudy/VennDiagram.png",
      width     = 4,
      height    = 4,
      units     = "in",
      res       = 200,
      pointsize = 8) # start export
      
grid::grid.newpage()
grid::grid.draw(VennDiagram::venn.diagram(list(satuRn=satuRn_results_comp6_top200,
                                               Limma_diffsplice=limmaRes_results_comp6_top200,
                                               DoubleExpSeq=DoubleExpSeq_results_comp6_top200), 
                                          filename = NULL, 
                                          main="Top200 DTU transcripts",
                                          main.fontface = "bold",
                                          cex = 2,
                                          main.cex = 2.5,
                                          cat.cex = 2,
                                          fill = c("black", "darkblue", "#56B4E9"),
                                          margin = 0.1,
                                          cat.just = list(c(1,-0.5),c(0.6,-0.5),c(1.3,0))))
dev.off()
```

