---
title: "4_Chen_S9S10.Rmd"
author: "Jeroen Gilis"
date: "08/12/2020"
output: html_document
---

This script (4_Chen_S9S10.Rmd) is a seperate script for reproducing figures S9 and S10 from our publication. **In order to run this script (4_Love_DTU_subset.Rmd), the dataset Love_benchmark_datasets_scaledTPM.Rdata (or, alternatively, Love_benchmark_datasets_count.Rdata) is required.** This file can either be generated with the 1_Love_prepare.Rmd script or downloaded from Zenodo.

```{r setup, include=FALSE, echo=FALSE}
require("knitr")
opts_knit$set(root.dir = rprojroot::find_rstudio_root_file())
```

# Load libraries

```{r,message=FALSE,warning=FALSE}
library(ggplot2)
library(iCOBRA)
```

# Load truth file

```{r}
load(file="./Data/Chen_benchmark_datasets_count.Rdata")
#load(file="./Data/Chen_benchmark_datasets_scaledTPM.Rdata")
metaInfo <- c(ChenBenchmarkLenient,ChenBenchmarkStringent)
rm(ChenBenchmarkLenient,ChenBenchmarkStringent)
```

# Load DTU performance results

```{r}
load(file="./Data/Chen_DTU_results_count.Rdata")
#load(file="./Data/Chen_DTU_results_scaledTPM.Rdata")

ChenBenchmark <- c(
    ChenDtuBenchmark_satuRn,
    ChenDtuBenchmark_limmaDiffsplice,
    ChenDtuBenchmark_DEXSeq,
    ChenDtuBenchmark_DRIMSeq,
    ChenDtuBenchmark_edgeRdiffsplice,
    ChenDtuBenchmark_DoubleExpSeq
)

### Remove empty entries (due to not tested - aka to many samples for reasonable runtime)
ChenBenchmark  <- ChenBenchmark[which(
    sapply(ChenBenchmark, function(x) ! is.null(x$dtuAnalysis)))]

rm(ChenDtuBenchmark_satuRn,
    ChenDtuBenchmark_limmaDiffsplice,
    ChenDtuBenchmark_DEXSeq,
    ChenDtuBenchmark_DRIMSeq,
    ChenDtuBenchmark_edgeRdiffsplice,
    ChenDtuBenchmark_DoubleExpSeq)
invisible(gc())
```

# Figure S9C

```{r}
tval <- ChenBenchmark[["satuRn_samples_used_50_rep_3_filterLenient"]]$dtuAnalysis$tvalue
pval <- ChenBenchmark[["satuRn_samples_used_50_rep_3_filterLenient"]]$dtuAnalysis$p_value_raw

zval <- qnorm(pval/2) * sign(tval)
zval_mid <- zval[abs(zval) < 10]
zval_mid <- zval_mid[!is.na(zval_mid)]

png("./Results/Chen_benchmark/figS9C.png",
    width     = 3.5,
    height    = 2.5,
    units     = "in",
    res       = 300,
    pointsize = 6) # start export
plot <- locfdr::locfdr(zval_mid, main = "Chen dataset - 50v50 - edgeR filter - repeat 3")
dev.off()
```

# Figure S9D

```{r}
dataset <- "satuRn_samples_used_50_rep_3_filterLenient"
Chen_select <- ChenBenchmark[[dataset]]$dtuAnalysis
truth <- metaInfo[[substr(dataset,8,nchar(dataset))]]$metaInfo

pvalues <- as.data.frame(cbind(Chen_select$p_value_raw,Chen_select$p_value))
colnames(pvalues) <- c("theoretical_null", "empirical_null")
rownames(pvalues) <- rownames(Chen_select)

rownames(truth) <- truth$TXNAME

cobra <- COBRAData(pval = pvalues, truth = truth)
cobra <- calculate_adjp(cobra)
cobra1perf <- calculate_performance(cobra, binary_truth = "txSwapped", cont_truth = "none", splv = "none", aspects = c("fdrtpr", "fdrtprcurve", "overlap"))

cobraplot <- prepare_data_for_plot(cobra1perf, colorscheme = "Dark2", facetted = TRUE)

new_col <- c(rep(c("black", "grey64", "#999999"),4),rep("white",6))
names(new_col) <- names(cobraplot@plotcolors)
cobraplot@plotcolors <- new_col

plot <- plot_fdrtprcurve(cobraplot, xaxisrange = c(0, 0.4), yaxisrange = c(0,1))

plot <- plot + 
    ggtitle(label="Chen dataset - 50v50 - edgeR filter - repeat 3") +
    xlab("FDP") +
    theme(strip.background = element_blank(),
    strip.text.x = element_blank(),
    plot.title = element_text(size=10),
    axis.text.x = element_text(size = 10),
    axis.text.y = element_text(size = 10),
    axis.title.x = element_text(size = 10),
    axis.title.y = element_text(size = 10))

png("./Results/Chen_benchmark/figS9D.png",
    width = 5.5,
    height = 4.5,
    units = "in",
    res = 200,
    pointsize = 4) # start export
print(plot)
dev.off()
```

# Figure S10A

```{r}
data <- ChenBenchmark[["DoubleExpSeq_samples_used_50_rep_3_filterLenient"]]$dtuAnalysis

png("./Results/Chen_benchmark/figS10A.png",
          width     = 3.5,
          height    = 2.5,
          units     = "in",
          res       = 300,
          pointsize = 6) # start export
hist(data$p_value,breaks=40,col="white",main = "Chen dataset - 50v50 - edgeR filter - repeat 3", xlab = "pvalues")
```

# Figure S10B

```{r}
direction <- data$sign
pval <- data$p_value

z <- qnorm(pval/2) * sign(direction)
z_working <- z[!is.na(z)]
z_working <- z_working[abs(z_working) < 10]

png("./Results/Chen_benchmark/figS10B.png",
        width     = 4.5,
        height    = 3,
        units     = "in",
        res       = 500,
        pointsize = 8)
current_plot <- suppressWarnings(locfdr::locfdr(z_working, main = "Chen dataset - 50v50 - edgeR filter - repeat 3"))
dev.off()
```




