---
title: "3_Gtex_visualize"
author: "Jeroen Gilis"
date: "02/11/2020"
output: html_document
---

**In order to run this script (3_Gtex_visualize.Rmd), the datasets 02_GTEx_benchmark_datasets_count.Rdata and 03_GTEx_DTU_results_count.Rdata are required.** These files can either be generated with the 1_Gtex_prepare.Rmd and 2_Gtex_DTU.Rmd scripts or downloaded from Zenodo. 

The results of this script (FDR-TPR curves) arealready available from our GitHub, under the directory satuRn_paper/Performance_benchmarks/Bulk_Gtex/Results/.

```{r setup, include=FALSE, echo=FALSE}
require("knitr")
opts_knit$set(root.dir = rprojroot::find_rstudio_root_file())
```

```{r,message=FALSE,warning=FALSE}
library(tidyverse)
library(iCOBRA)
```

# Load benchmark datasets (contains the ground truth)

As generated in script 1_Gtex_prepare.Rmd

```{r}
### Load and concatenate data
load(file="./Data/02_GTEx_benchmark_datasets_count.Rdata")
#load(file="01_gtex_benchmark_data_2017_scaledTPM.Rdata") # TPM equivalent
metaInfo <- c(gtexBenchmarkDataLenient,gtexBenchmarkDataStringent)
rm(gtexBenchmarkDataLenient,gtexBenchmarkDataStringent)
invisible(gc())
```

# Load DTU analysis results

As generated in script 2_Gtex_DTU.Rmd

```{r}
load(file="./Data/03_GTEx_DTU_results_count.Rdata")
#load(file="02_Gtex_DTU_analysis_2017_scaledTPM_full.Rdata") # TPM equivalent

gtexBenchmark <- c(
    gtexDtuBenchmark_DEXSeq,
    gtexDtuBenchmark_DoubleExpSeq,
    gtexDtuBenchmark_DRIMSeq,
    gtexDtuBenchmark_edgeRdiffsplice,
    gtexDtuBenchmark_limmaDiffsplice,
    gtexDtuBenchmark_NBSplice,
    gtexDtuBenchmark_satuRn
)

### Remove empty enteries (due to not tested due to too many samples for reasonable runtime)
gtexBenchmark  <- gtexBenchmark[which(
    sapply(gtexBenchmark, function(x) ! is.null(x$dtuAnalysis))
)]

rm(gtexDtuBenchmark_DEXSeq,
    gtexDtuBenchmark_DRIMSeq,
    gtexDtuBenchmark_DoubleExpSeq,
    gtexDtuBenchmark_edgeRdiffsplice,
    gtexDtuBenchmark_limmaDiffsplice,
    gtexDtuBenchmark_NBSplice,
    gtexDtuBenchmark_satuRn)
invisible(gc())
```

# Helper function 
To get pvalues for the different methods on the different datasets

```{r}
get_pvalues <- function(dataset,nrep,nmethods) {
  
  get_all_txs <- function(dataset){
  
    all_txs <- c()
    current_txs <- as.character(dataset[[1]]$dtuAnalysis$TXNAME)
    all_txs <- c(all_txs, current_txs)
    all_txs <- unique(all_txs)
    return(all_txs)
  }

  all_txs <- get_all_txs(dataset)
  nmethods <- length(dataset)
  pvalues <- matrix(data=NA, nrow=length(all_txs), ncol=nmethods)
  rownames(pvalues) <- all_txs
  
  for (i in 1:nmethods) {
    pvalues_new <- data.frame(dataset[[i]]$dtuAnalysis$p_value)
    rownames(pvalues_new) <- as.character(dataset[[i]]$dtuAnalysis$TXNAME)
    colnames(pvalues_new) <- names(dataset)[i]

    pvalues[,i] <- pvalues_new[match(rownames(pvalues),rownames(pvalues_new)),1]
  }
  
  pvalues <- apply(pvalues, 2, as.numeric)
  rownames(pvalues) <- all_txs
  pvalues <- as.data.frame(pvalues)
  pvalues <- pvalues[order(rownames(pvalues)),]
  colnames(pvalues) <- methods

  return(pvalues)
}
```

# Visualization function

```{r}
visualize_datasets <- function(dataset,metaInfo,nFilters,nSampleSizes,methods,colors,titles,selection) {
    nPlots <- nFilters * nSampleSizes
    print(paste("This function will generate",  nPlots, "plots"))
  
    # for each of the requested benchmarks, do:
    for (j in 1:nPlots) {
        
        print(j) # print progress
        gtexBenchmark_current_outer <- dataset[grepl(selection[j], names(dataset))] # select the requested analyses
        
        selected_methods <- c()
        for (i in 1:length(methods)) {
            selected_methods <- c(selected_methods,grep(methods[i], names(gtexBenchmark_current_outer)))
        }
        
        gtexBenchmark_current_outer <- gtexBenchmark_current_outer[selected_methods] # name ordered

        metaInfo_current_outer <- metaInfo[grepl(selection[j], names(metaInfo))]

        resList <- c() # to store results of each repeat later
        #nRep=c("rep_1_","rep_2","rep_3","rep_4","rep_5","rep_6","rep_7","rep_8","rep_9","rep_10") # i.e. we repeated the 5v5 edgeR filter analysis three times
        nRep=c("rep_1_","rep_2","rep_3")
        # for each of k repeats, do:
        for (k in 1:length(nRep)) {
            gtexBenchmark_current_inner <- gtexBenchmark_current_outer[grepl(nRep[k], names(gtexBenchmark_current_outer))] # get analysis data of k-th repeat
            metaInfo_current_inner <- metaInfo_current_outer[grepl(nRep[k], names(metaInfo_current_outer))] # get truth data of k-th repeat

            pvalues <- get_pvalues(gtexBenchmark_current_inner,methods)

            truth_file <- gtexBenchmark_current_inner[[1]]$dtuAnalysis
            txSwapped <- metaInfo_current_inner[1] # txSwapped is that of the first repeat but is the same for all (3) repeats
            truth_file <- truth_file[,c("TXNAME","p_value")]
    
            txSwapped[[1]]$metaInfo <- txSwapped[[1]]$metaInfo[match(truth_file$TXNAME,txSwapped[[1]]$metaInfo$TXNAME),]
    
            truth_file$gene_modified <- as.numeric(txSwapped[[1]]$metaInfo$txSwapped)
            truth_file <- truth_file[order(truth_file$TXNAME),]
            rownames(truth_file) <- truth_file$TXNAME

            resList[[k]] <- list(pvalues,truth_file)
        }
        
        # concatenate the results of the repeated analyses into one dataframe in order to calculate average performances over repeats (required for iCobra). Avoid repeated rownames. 
        rownames(resList[[1]][[1]]) <- paste0(rownames(resList[[1]][[1]]), "_1")
        rownames(resList[[2]][[1]]) <- paste0(rownames(resList[[2]][[1]]), "_2")
        rownames(resList[[3]][[1]]) <- paste0(rownames(resList[[3]][[1]]), "_3")

        pvalues_full <- rbind(resList[[1]][[1]],resList[[2]][[1]],resList[[3]][[1]])
        
        # concatenate the metadata of the repeated analyses into one dataframe in order to calculate average performances over repeats (required for iCobra). Avoid repeated rownames. 
        rownames(resList[[1]][[2]]) <- paste0(rownames(resList[[1]][[2]]), "_1")
        rownames(resList[[2]][[2]]) <- paste0(rownames(resList[[2]][[2]]), "_2")
        rownames(resList[[3]][[2]]) <- paste0(rownames(resList[[3]][[2]]), "_3")

        truth_file_full <- rbind(resList[[1]][[2]],resList[[2]][[2]],resList[[3]][[2]])
        
        # generate plot with iCobra
        cobra <- COBRAData(pval = pvalues_full, truth = truth_file_full)
        cobra <- calculate_adjp(cobra)
        cobra1perf <- calculate_performance(cobra, binary_truth = "gene_modified", cont_truth = "none", splv = "none", aspects = c("fdrtpr", "fdrtprcurve", "overlap"))
        cobraplot <- prepare_data_for_plot(cobra1perf, colorscheme = "Dark2", facetted = TRUE)
        
        # customize colors
        new_col <- colors[which(names(colors)%in%methods)]
        new_col <- c(rep(c(new_col, "#999999"),4),rep("white",2*(length(methods)+1)))
        names(new_col) <- names(cobraplot@plotcolors)
        cobraplot@plotcolors <- new_col

        # create plot
        plot <- plot_fdrtprcurve(cobraplot, xaxisrange = c(0, 0.4), yaxisrange = c(0,1))
        plot <- plot +
            ggtitle(titles[j]) +
            theme(strip.background = element_blank(),
            strip.text.x = element_blank(),
            plot.title = element_text(size=10),
            axis.text.x = element_text(size = 10),
            axis.text.y = element_text(size = 10),
            axis.title.x = element_text(size = 10),
            axis.title.y = element_text(size = 10))
        
        # save plot
        png(paste0("./Performance_benchmarks/Bulk_Gtex/Results/FDRTPR_", gsub(" ", "_", titles[j]), ".png"),
        width     = 5.5,
        height    = 4.5,
        units     = "in",
        res       = 200,
        pointsize = 4) # start export
        print(plot)
        dev.off()
  }
}
```

# Visualize the results

```{r}
selection <- c("5_rep_.(_|0_)filterLenient","20_rep_.(_|0_)filterLenient","5_rep_.(_|0_)filterStringent","20_rep_.(_|0_)filterStringent")

colors<- c("#E69F00", "#56B4E9", "#009E73", "#F0E442", "#0072B2", "#D55E00", "black")
names(colors) <- c("DEXSeq","DoubleExpSeq","DRIMSeq","edgeR_diffsplice", "limma_diffsplice" ,"NBSplice","satuRn")

methods <- c("DEXSeq","DoubleExpSeq","DRIMSeq","edgeR_diffsplice", "limma_diffsplice" ,"NBSplice","satuRn")

titles <- c("5 versus 5 - edgeR filter - count","20 versus 20 - edgeR filter - count","5 versus 5 - DRIMSeq filter - count","20 versus 20 - DRIMSeq filter - count")

visualize_datasets(dataset=gtexBenchmark,
                   metaInfo=metaInfo,
                   nFilters=2,
                   nSampleSizes=2,
                   methods=methods,
                   colors=colors,
                   titles=titles,
                   selection=selection)

selection <- c("50_rep_.(_|0_)filterLenient","50_rep_.(_|0_)filterStringent")

colors<- c("#56B4E9", "#F0E442", "#0072B2", "black")
names(colors) <- c("DoubleExpSeq","edgeR_diffsplice", "limma_diffsplice","satuRn")

methods <- c("DoubleExpSeq","edgeR_diffsplice", "limma_diffsplice","satuRn")

titles <- c("50 versus 50 - edgeR filter - count","50 versus 50 - DRIMSeq filter - count")

visualize_datasets(dataset=gtexBenchmark,
                   metaInfo=metaInfo,
                   nFilters=2,
                   nSampleSizes=1,
                   methods=methods,
                   colors=colors,
                   titles=titles,
                   selection=selection)
```




